---
title: "Arch and Func Interps"
author: "Rudolf Cesaretti"
date: "2025-03-06"
output: html_document
---



# ArchFuncInterps



```{r}
rm(list = ls())

```

# Load Packages

```{r}
# Load the required package
library(tidyverse)
library(readxl)
library(stringi)
library(arrow)

```


# Set Working Directory

```{r}
DIR = "C:/Users/TJ McMote/ASU Dropbox/Rudolf Cesaretti/GitHubRepos/TMP-database/"
```



```{r}
DF9b_data_char <- read_csv(paste0(DIR, "data/raw/DF9b_data_char.csv"), locale = locale(encoding = "Latin1")) 
```
```{r}
archInterp <- DF9b_data_char %>% 
  select(SSN, arch1PMic, arch1McTl, arch1XlMe, arch1Oxto, arch2PMic, arch2McTl, 
         arch2XlMe, arch2Oxto, func1PMic, func1McTl, func1XlMe, func1Oxto, 
         func2PMic, func2McTl, func2XlMe, func2Oxto)

colnames(archInterp) <- c("SSN", "archInterp1_PMic", "archInterp1_McTl", "archInterp1_XlMe", "archInterp1_Oxto", "archInterp2_PMic", "archInterp2_McTl", "archInterp2_XlMe", "archInterp2_Oxto", "funcInterp1_PMic", "funcInterp1_McTl", "funcInterp1_XlMe", "funcInterp1_Oxto", "funcInterp2_PMic", "funcInterp2_McTl", "funcInterp2_XlMe", "funcInterp2_Oxto")

archInterp_long <- archInterp %>%
  pivot_longer(
    cols = -SSN,  # Keep SSN as an identifier column
    names_to = c(".value", "Period"),  # Extract variable names and period
    names_sep = "_"
  )

archInterp_long2 <- archInterp_long %>% 
  rowid_to_column(var = "ID") %>% 
  mutate(Period = ifelse(Period == "PMic", "EC", 
                         ifelse(Period == "McTl", "MC", 
                                ifelse(Period == "XlMe", "LC", "ET"))))


archInterp_long3 <- archInterp_long2 %>% 
  mutate(count = 1) %>%
  group_by(archInterp1, archInterp2, funcInterp1, funcInterp2) %>%
  summarize(count = sum(count, na.rm = T))
  
archInterp_long4 <- archInterp_long2 %>% 
  mutate(count = 1) %>%
  group_by(archInterp1, funcInterp1) %>%
  summarize(count = sum(count, na.rm = T))

archInterp_long5 <- archInterp_long2 %>% 
  mutate(count = 1) %>%
  group_by(archInterp2, funcInterp2) %>%
  summarize(count = sum(count, na.rm = T))
```

SSN, arch1PMic, arch1McTl, arch1XlMe, arch1Oxto, arch2PMic, arch2McTl, arch2XlMe, arch2Oxto, func1PMic, func1McTl, func1XlMe, func1Oxto, func2PMic, func2McTl, func2XlMe, func2Oxto


SSN, arch1_PMic, arch1_McTl, arch1_XlMe, arch1_Oxto, arch2_PMic, arch2_McTl, arch2_XlMe, arch2_Oxto, func1_PMic, func1_McTl, func1_XlMe, func1_Oxto, func2_PMic, func2_McTl, func2_XlMe, func2_Oxto

"SSN", "archInterp1_PMic", "archInterp1_McTl", "archInterp1_XlMe", "archInterp1_Oxto", "archInterp2_PMic", "archInterp2_McTl", "archInterp2_XlMe", "archInterp2_Oxto", "funcInterp1_PMic", "funcInterp1_McTl", "funcInterp1_XlMe", "funcInterp1_Oxto", "funcInterp2_PMic", "funcInterp2_McTl", "funcInterp2_XlMe", "funcInterp2_Oxto"

```{r}
write.csv(archInterp_long3, paste0(DIR, "data/raw/ArchFuncInterps1.csv"), row.names = F, fileEncoding = "Latin1")

write.csv(archInterp_long4, paste0(DIR, "data/raw/ArchFuncInterps2.csv"), row.names = F, fileEncoding = "Latin1")

write.csv(archInterp_long5, paste0(DIR, "data/raw/ArchFuncInterps3.csv"), row.names = F, fileEncoding = "Latin1")


```


