---
title: "Arch and Func Interps"
author: "Rudolf Cesaretti"
date: "2025-03-06"
output: html_document
---



# ArchFuncInterps



```{r}
rm(list = ls())

```

# Load Packages

```{r}
# Load the required package
library(tidyverse)
library(readxl)
library(stringi)
library(arrow)

```


# Set Working Directory

```{r}
DIR = "C:/Users/TJ McMote/ASU Dropbox/Rudolf Cesaretti/GitHubRepos/TMP-database/"
```



```{r}
DF9b_data_char <- read_csv(paste0(DIR, "data/raw/DF9b_data_char.csv"), locale = locale(encoding = "Latin1")) 
```




```{r}
archInterp <- DF9b_data_char %>% 
  select(SSN, arch1PMic, arch1McTl, arch1XlMe, arch1Oxto, arch2PMic, arch2McTl, 
         arch2XlMe, arch2Oxto, func1PMic, func1McTl, func1XlMe, func1Oxto, 
         func2PMic, func2McTl, func2XlMe, func2Oxto)

colnames(archInterp) <- c("SSN", "archInterp1_PMic", "archInterp1_McTl", "archInterp1_XlMe", "archInterp1_Oxto", "archInterp2_PMic", "archInterp2_McTl", "archInterp2_XlMe", "archInterp2_Oxto", "funcInterp1_PMic", "funcInterp1_McTl", "funcInterp1_XlMe", "funcInterp1_Oxto", "funcInterp2_PMic", "funcInterp2_McTl", "funcInterp2_XlMe", "funcInterp2_Oxto")

archInterp_long <- archInterp %>%
  pivot_longer(
    cols = -SSN,  # Keep SSN as an identifier column
    names_to = c(".value", "PhaseAbbr"),  # Extract variable names and period
    names_sep = "_"
  ) %>% 
  rowid_to_column(var = "ID") %>% 
  mutate(Period = ifelse(PhaseAbbr == "PMic", "EC", 
                         ifelse(PhaseAbbr == "McTl", "MC", 
                                ifelse(PhaseAbbr == "XlMe", "LC", "ET"))),
         Phase = ifelse(PhaseAbbr == "PMic", "Patlachique-Tzacualli", 
                         ifelse(PhaseAbbr == "McTl", "Miccaotli-Tlamimilolpa", 
                                ifelse(PhaseAbbr == "XlMe", "Xolalpan-Metepec", "Oxtoticpac")))) %>% 
  select(ID,SSN,Phase,PhaseAbbr,Period,archInterp1,funcInterp1,archInterp2,funcInterp2) %>% 
  rename(
    archInterp_1 = archInterp1,
    archInterp_2 = archInterp2,
    funcInterp_1 = funcInterp1,
    funcInterp_2 = funcInterp2
  )




archInterp_long2 <- archInterp_long %>% 
  pivot_longer(
    cols = c(archInterp_1, archInterp_2, funcInterp_1, funcInterp_2),
    names_to = c(".value", "InterpNum"),
    names_sep = "_"
  ) %>% 
  mutate(
    InterpNum = as.numeric(InterpNum),
    Interp = ifelse(InterpNum == 1, "Primary", "Secondary")
  ) %>% 
  select(ID,SSN,Phase,PhaseAbbr,Period,Interp,InterpNum,archInterp,funcInterp)

archInterp_summary1 <- archInterp_long2 %>% 
  mutate(count = 1) %>%
  group_by(archInterp, funcInterp) %>%
  summarize(count = sum(count, na.rm = T)) %>% 
  ungroup() %>%
  group_by(archInterp) %>%
  mutate(arch_total = sum(count, na.rm = T),
         arch_pct = count/arch_total) %>% 
  ungroup() %>%
  group_by(funcInterp) %>%
  mutate(func_total = sum(count, na.rm = T),
         func_pct = count/func_total) %>%
  ungroup() %>%
  arrange(funcInterp, -count) %>%
  rowid_to_column(var = "ORD_F") %>% 
  arrange(archInterp, -count) %>%
  rowid_to_column(var = "ORD_A")

archInterp_summary1b <- archInterp_summary1 %>% 
  mutate(PairIndex = arch_pct + func_pct) %>% 
  arrange(-PairIndex)

archInterp_summary1_LC <- archInterp_long2 %>% 
  filter(Period == "LC") %>%
  mutate(count = 1) %>%
  group_by(archInterp, funcInterp) %>%
  summarize(count = sum(count, na.rm = T)) %>% 
  ungroup() %>%
  group_by(archInterp) %>%
  mutate(arch_total = sum(count, na.rm = T),
         arch_pct = count/arch_total) %>% 
  ungroup() %>%
  group_by(funcInterp) %>%
  mutate(func_total = sum(count, na.rm = T),
         func_pct = count/func_total) %>%
  ungroup() %>%
  arrange(funcInterp, -count) %>%
  rowid_to_column(var = "ORD_F") %>% 
  arrange(archInterp, -count) %>%
  rowid_to_column(var = "ORD_A")

archInterp_summary1b_LC <- archInterp_summary1_LC %>% 
  mutate(PairIndex = arch_pct + func_pct) %>% 
  arrange(-PairIndex)

join <- archInterp_summary1b_LC %>% 
  mutate(JOIN = paste(archInterp, funcInterp, sep = "; ")) %>% 
  select(JOIN, count, arch_pct, func_pct, PairIndex) %>%
  rename(count_LC = count, arch_pct_LC = arch_pct, func_pct_LC = func_pct, PairIndex_LC = PairIndex)


archInterp_summary1c <- archInterp_summary1b %>% 
  mutate(JOIN = paste(archInterp, funcInterp, sep = "; ")) %>% 
  left_join(join, by = "JOIN") %>% 
  select(-JOIN)




archInterp_summary2 <- archInterp_long2 %>% 
  filter(Period == "LC") %>%
  mutate(count = 1) %>%
  group_by(archInterp, funcInterp) %>%
  summarize(count = sum(count, na.rm = T))

write.csv(archInterp_summary1c, paste0(DIR, "data/raw/archInterp_summary1c.csv"), row.names = F, fileEncoding = "Latin1")

```






```{r}

archInterp_long3 <- archInterp_long2 %>% 
  mutate(count = 1) %>%
  group_by(archInterp1, archInterp2, funcInterp1, funcInterp2) %>%
  summarize(count = sum(count, na.rm = T))
  
archInterp_long4 <- archInterp_long2 %>% 
  mutate(count = 1) %>%
  group_by(archInterp1, funcInterp1) %>%
  summarize(count = sum(count, na.rm = T))

archInterp_long5 <- archInterp_long2 %>% 
  mutate(count = 1) %>%
  group_by(archInterp2, funcInterp2) %>%
  summarize(count = sum(count, na.rm = T))
```



SSN, arch1PMic, arch1McTl, arch1XlMe, arch1Oxto, arch2PMic, arch2McTl, arch2XlMe, arch2Oxto, func1PMic, func1McTl, func1XlMe, func1Oxto, func2PMic, func2McTl, func2XlMe, func2Oxto


SSN, arch1_PMic, arch1_McTl, arch1_XlMe, arch1_Oxto, arch2_PMic, arch2_McTl, arch2_XlMe, arch2_Oxto, func1_PMic, func1_McTl, func1_XlMe, func1_Oxto, func2_PMic, func2_McTl, func2_XlMe, func2_Oxto

"SSN", "archInterp1_PMic", "archInterp1_McTl", "archInterp1_XlMe", "archInterp1_Oxto", "archInterp2_PMic", "archInterp2_McTl", "archInterp2_XlMe", "archInterp2_Oxto", "funcInterp1_PMic", "funcInterp1_McTl", "funcInterp1_XlMe", "funcInterp1_Oxto", "funcInterp2_PMic", "funcInterp2_McTl", "funcInterp2_XlMe", "funcInterp2_Oxto"

```{r}
write.csv(archInterp_long3, paste0(DIR, "data/raw/ArchFuncInterps1.csv"), row.names = F, fileEncoding = "Latin1")

write.csv(archInterp_long4, paste0(DIR, "data/raw/ArchFuncInterps2.csv"), row.names = F, fileEncoding = "Latin1")

write.csv(archInterp_long5, paste0(DIR, "data/raw/ArchFuncInterps3.csv"), row.names = F, fileEncoding = "Latin1")


```














cultivated field
residence of uncertain status
unoccupied, not clear if cultivated or not
platform of probable religious significance
platform with other structures on it
# Occupational Vars

```{r}
OccuVarsCalc = archInterp_long %>% 
  mutate(xxx = ifelse(is.na(archInterp_1) & !is.na(funcInterp_1), T, F),
         archInterp_1 = case_when(
           xxx == T & funcInterp_1 == "cultivated field" ~ "probably not occupied or utilized in this period",
           xxx == T & funcInterp_1 == "unoccupied, not clear if cultivated or not" ~ "probably not occupied or utilized in this period",
           xxx == T & funcInterp_1 == "residence of uncertain status" ~ "room group",
           xxx == T & funcInterp_1 == "platform of probable religious significance" ~ "platform with other structures on it",
           .default = archInterp_1),
         Occu_1 = archInterp_1,
         Occu_2 = archInterp_2,
         Occu_2 = case_when(
           Occu_2 == "no plausible alternative to the primary interpretation" ~ Occu_1,
           Occu_2 == "primary interpretation is the best assessment of a very ambiguous situation" ~ Occu_1,
           .default = Occu_2
         ),
         Occu_2 = case_when(
           Occu_2 == "room group" ~ "2",
           Occu_2 == "casa humilde" ~ "2",
           Occu_2 == "apartment compound" ~ "2",
           Occu_2 == "temple base (incl. platforms & multi-tiered platforms)" ~ "2",
           Occu_2 == "pyramid (more than ca. 10m high)" ~ "2",
           Occu_2 == "simple platform" ~ "2",
           Occu_2 == "platform with other structures on it" ~ "2",
           Occu_2 == "platform in plaza (momoztli)" ~ "1",
           Occu_2 == "plaza" ~ "1",
           Occu_2 == "free-standing wall or segment of wall" ~ "1",
           Occu_2 == "access-way (alley, street, avenue, etc.)" ~ "1",
           Occu_2 == "avenue divider or possible avenue divider" ~ "1",
           Occu_2 == "cemetery" ~ "1",
           Occu_2 == "dam" ~ "1",
           Occu_2 == "occupation without durable structures" ~ "1",
           Occu_2 == "probably not occupied or utilized in this period" ~ "0",
           Occu_2 == "cave" ~ "0",
           Occu_2 == "other" ~ "1"),
         Occu_2 = as.numeric(Occu_2),
         Occu_1 = case_when(
           Occu_1 == "room group" ~ "2",
           Occu_1 == "casa humilde" ~ "2",
           Occu_1 == "apartment compound" ~ "2",
           Occu_1 == "temple base (incl. platforms & multi-tiered platforms)" ~ "2",
           Occu_1 == "pyramid (more than ca. 10m high)" ~ "2",
           Occu_1 == "simple platform" ~ "2",
           Occu_1 == "platform with other structures on it" ~ "2",
           Occu_1 == "platform in plaza (momoztli)" ~ "1",
           Occu_1 == "plaza" ~ "1",
           Occu_1 == "free-standing wall or segment of wall" ~ "1",
           Occu_1 == "access-way (alley, street, avenue, etc.)" ~ "1",
           Occu_1 == "avenue divider or possible avenue divider" ~ "1",
           Occu_1 == "cemetery" ~ "1",
           Occu_1 == "dam" ~ "1",
           Occu_1 == "occupation without durable structures" ~ "1",
           Occu_1 == "probably not occupied or utilized in this period" ~ "0",
           Occu_1 == "cave" ~ "0",
           Occu_1 == "other" ~ "1"),
         Occu_1 = as.numeric(Occu_1),
         #Occu = if_else(is.na(Occu_1) & is.na(Occu_2),
         #                  NA_real_,
         #                  coalesce(Occu_1, 0) + coalesce(Occu_2, 0)),
         Occu = case_when(
           is.na(Occu_1) & is.na(Occu_2) ~ NA_character_,
           is.na(Occu_1) & Occu_2 == 0 ~ "0. No evidence for occupation",
           is.na(Occu_1) & Occu_2 == 1 ~ "1. Likely either less intensive or archaeologically ephemeral occupation",
           is.na(Occu_1) & Occu_2 == 2 ~ "2. Evidence for permanent occupation of some form",
           Occu_1 == 0 & is.na(Occu_2) ~ "0. No evidence for occupation",
           Occu_1 == 0 & Occu_2 == 0 ~ "0. No evidence for occupation",
           Occu_1 == 0 & Occu_2 == 1 ~ "1. Likely either less intensive or archaeologically ephemeral occupation",
           Occu_1 == 0 & Occu_2 == 2 ~ "1. Likely either less intensive or archaeologically ephemeral occupation",
           Occu_1 == 1 & is.na(Occu_2) ~ "1. Likely either less intensive or archaeologically ephemeral occupation",
           Occu_1 == 1 & Occu_2 == 0 ~ "1. Likely either less intensive or archaeologically ephemeral occupation",
           Occu_1 == 1 & Occu_2 == 1 ~ "1. Likely either less intensive or archaeologically ephemeral occupation",
           Occu_1 == 1 & Occu_2 == 2 ~ "2. Evidence for permanent occupation of some form",
           Occu_1 == 2 & is.na(Occu_2) ~ "3. Permanent occupation likely in the form of domestic structures or public buildings",
           Occu_1 == 2 & Occu_2 == 0 ~ "2. Evidence for permanent occupation of some form",
           Occu_1 == 2 & Occu_2 == 1 ~ "3. Permanent occupation likely in the form of domestic structures or public buildings",
           Occu_1 == 2 & Occu_2 == 2 ~ "3. Permanent occupation likely in the form of domestic structures or public buildings"))


OccuVars <- OccuVarsCalc %>% 
  pivot_wider(
    id_cols = SSN,
    names_from = Period,
    names_prefix = "Occu_",
    values_from = Occu
  )

write.csv(OccuVarsCalc, paste0(DIR, "data/raw/OccuVarsCalc.csv"), row.names = F, fileEncoding = "Latin1")
write.csv(OccuVars, paste0(DIR, "data/raw/OccuVars.csv"), row.names = F, fileEncoding = "Latin1")
  
```

"totAll"               "totPrec"              "totCuan"              "totTezo"              ""              ""             
[211] "totMicc"              "totTlam"              "totXola"              "totMete"              "totOxto"              "totXome"             
[217] "totMaza"              "totAzte"


```{r}
cer = DF9b_data_char[,c(1,194,204:218,290)]

cer=cer %>% 
  mutate(
    cerAbundNum = case_when(
      ceramicAbundance == "Absent" ~ 0,
      ceramicAbundance == "Very Sparse" ~ 1,
      ceramicAbundance == "Very Sparse to Sparse" ~ 2,
      ceramicAbundance == "Sparse" ~ 3,
      ceramicAbundance == "Sparse to Moderate" ~ 4,
      ceramicAbundance == "Moderate" ~ 5,
      ceramicAbundance == "Moderate to Abundant" ~ 6,
      ceramicAbundance == "Abundant" ~ 7,
      ceramicAbundance == "Abundant to Very Abundant" ~ 8,
      ceramicAbundance == "Very Abundant" ~ 9,
      TRUE ~ NA_real_),
    cerAbund_DensEst = case_when(
      ceramicAbundance == "Absent" ~ 0,
      ceramicAbundance == "Very Sparse" ~ 2,
      ceramicAbundance == "Very Sparse to Sparse" ~ 5,
      ceramicAbundance == "Sparse" ~ 10,
      ceramicAbundance == "Sparse to Moderate" ~ 25,
      ceramicAbundance == "Moderate" ~ 50,
      ceramicAbundance == "Moderate to Abundant" ~ 100,
      ceramicAbundance == "Abundant" ~ 200,
      ceramicAbundance == "Abundant to Very Abundant" ~ 350,
      ceramicAbundance == "Very Abundant" ~ 500,
      TRUE ~ NA_real_),
    areaSite= areaSite*10,
    cerCollectAreaDens = totAll/areaSite,
    EC = totPatl + totTzac,
    MC = totMicc + totTlam,
    LC = totXola + totMete,
    ET = totOxto + totXome,
    pctEC = EC/totAll,
    pctMC = MC/totAll,
    pctLC = LC/totAll,
    pctET = ET/totAll,
    densEC = cerAbund_DensEst*pctEC,
    densMC = cerAbund_DensEst*pctMC,
    densLC = cerAbund_DensEst*pctLC,
    densET = cerAbund_DensEst*pctET,
    areaDensEC = EC/areaSite,
    areaDensMC = MC/areaSite,
    areaDensLC = LC/areaSite,
    areaDensET = ET/areaSite)

OccuVars2 = OccuVars %>% 
  mutate(
    OccuNum_EC = case_when(
      Occu_EC == "0. No evidence for occupation" ~ 0,
      Occu_EC == "1. Likely either less intensive or archaeologically ephemeral occupation" ~ 1,
      Occu_EC == "2. Evidence for permanent occupation of some form" ~ 2,
      Occu_EC == "3. Permanent occupation likely in the form of domestic structures or public buildings" ~ 3),
    OccuNum_MC = case_when(
      Occu_MC == "0. No evidence for occupation" ~ 0,
      Occu_MC == "1. Likely either less intensive or archaeologically ephemeral occupation" ~ 1,
      Occu_MC == "2. Evidence for permanent occupation of some form" ~ 2,
      Occu_MC == "3. Permanent occupation likely in the form of domestic structures or public buildings" ~ 3),
    OccuNum_LC = case_when(
      Occu_LC == "0. No evidence for occupation" ~ 0,
      Occu_LC == "1. Likely either less intensive or archaeologically ephemeral occupation" ~ 1,
      Occu_LC == "2. Evidence for permanent occupation of some form" ~ 2,
      Occu_LC == "3. Permanent occupation likely in the form of domestic structures or public buildings" ~ 3),
    OccuNum_ET = case_when(
      Occu_ET == "0. No evidence for occupation" ~ 0,
      Occu_ET == "1. Likely either less intensive or archaeologically ephemeral occupation" ~ 1,
      Occu_ET == "2. Evidence for permanent occupation of some form" ~ 2,
      Occu_ET == "3. Permanent occupation likely in the form of domestic structures or public buildings" ~ 3),
    OccuNum_Total = OccuNum_EC + OccuNum_MC + OccuNum_LC + OccuNum_ET,
    OccuNum_Max = pmax(OccuNum_EC, OccuNum_MC, OccuNum_LC, OccuNum_ET),
    OccuBool_EC = ifelse(OccuNum_EC > 0, 1, 0),
    OccuBool_MC = ifelse(OccuNum_MC > 0, 1, 0),
    OccuBool_LC = ifelse(OccuNum_LC > 0, 1, 0),
    OccuBool_ET = ifelse(OccuNum_ET > 0, 1, 0),
    OccuBool_Total = OccuBool_EC + OccuBool_MC + OccuBool_LC + OccuBool_ET) %>% 
  left_join(cer, by = "SSN")

    "1. Likely either less intensive or archaeologically ephemeral occupation"
plot(OccuVars2$OccuBool_ET, OccuVars2$ET)
> plot(log(cer$cerCollectAreaDens), cer$cerAbund_DensEst)
> plot(log(cer$cerCollectAreaDens), log(cer$cerAbund_DensEst))
```


