---
title: "DF8 Wide"
author: "Rudolf Cesaretti"
date: "2025-01-21"
output: html_document
---


# Load Packages

```{r}
# Load the required package
library(tidyverse)
library(readxl)
library(stringi)

```


# Set Working Directory

```{r}
DIR = "C:/Users/TJ McMote/ASU Dropbox/Rudolf Cesaretti/GitHubRepos/TMP-database/"
```


# Create DF8 Wide 

## Merge DF8 Tables for Raw (original numeric) Values Dataset


### Step 1: Import all excel files from the specified directory into a list of dataframes

```{r}

file_list <- list.files(
  path = paste0(DIR, "data/raw/DF8"),
  pattern = "\\.xlsx$", full.names = TRUE
)
file_names <- tools::file_path_sans_ext(basename(file_list))
data_list <- map(file_list, read_excel)
names(data_list) <- file_names

```

### Step 2: Rearrange the list so the last dataframe is first, then join all
```{r}

# Rearrange the list so the last dataframe is first, then join all
DF8_wide <- data_list %>%
  {c(tail(., 1), head(., -1))} %>%  # Reorder the list
  reduce(full_join, by = "SSN") %>% # Join all dataframes by SSN
  rename(`SITE_` = `Site.`,
         `UNIT_` = `Unit`,
         `DRAINS_ARCH` = `DRAINS.x`,
         `DRAINS_MODERN` = `DRAINS.y`) # Rename columns

```

### Step 3: Check for any discrepancies between the SITE and UNIT columns

```{r}

checksum <- DF8_wide %>% 
  dplyr::select(SSN, SITE, SITE_, UNIT, UNIT_) %>% 
  mutate(check_SITE = ifelse(SITE == SITE_, 1, 0),
         check_UNIT = ifelse(UNIT == UNIT_, 1, 0)) %>% 
  summarize(checksum_SITE = sum(check_SITE),
            checksum_UNIT = sum(check_UNIT))

if (checksum$checksum_SITE == nrow(DF8_wide) & checksum$checksum_UNIT == nrow(DF8_wide)) {
  print("CHECKS PASSED: SITE + SITE_ and UNIT + UNIT_ columns are identical across all rows")
  DF8_wide <- DF8_wide %>% 
    dplyr::select(-SITE, -UNIT_) %>% 
    rename(SITE = SITE_)
} else {
  print("WARNING: SITE + SITE_ and UNIT + UNIT_ columns are NOT identical across all rows")
}

```

### Step 4: Save the DF8_wide dataframe to a CSV file

```{r}
write_csv(DF8_wide, paste0(DIR, "data/raw/DF8_wide.csv"))
```



## Process DF8 wide dataset values for Standardized Dataset








