---
title: "DF9 Wide"
author: "Rudolf Cesaretti"
date: "2025-01-21"
output: html_document
---



# Create DF9 Wide for Cross-Checking the Data


### Step 0: Import all excel files from the specified directory

```{r}

file_list <- list.files(
  path = paste0(DIR, "data/raw/DF9"),
  pattern = "\\.xlsx$", full.names = TRUE
)
file_names <- tools::file_path_sans_ext(basename(file_list))
data_list <- map(file_list, read_excel)
names(data_list) <- file_names
#list2env(data_list, envir = .GlobalEnv)

```
"admin" is the base table

-- 5050-row tables --

WorkingHierarchy
`a_FuncInterp1 for digitization`
archInterp
archMaterial




```{r}

ceramics = cerPhTot %>% 
  left_join(cerVessel, by = "ID") %>% 
  left_join(cerNonVessel, by = "ID")

lithics = lithicFlaked %>% 
  left_join(lithicGround, by = "ID")

artifactOther

```






complexData
complexMacroData

condition
description

location

Plazas
workshop



materials
materialsLabel


z_archInterp1
z_archInterp2
z_funcInterp1
z_funcInterp2
z_vegetation
HierarchyCode



personnelLabel
"personnel error check"
fieldWorkers
labAnalysts


```{r}




colz <- c("fieldWorker1", "fieldWorker2", "fieldWorker3", "fieldWorker4", "fieldWorker5",
                    "ceramicAnalyst1", "ceramicAnalyst2", "ceramicAnalyst3")


personnel <- personnelLabel %>% 
  mutate(FullName = paste(first_name, last_name, sep = " ")) %>% 
  select(-last_name, -first_name) %>% 
  mutate(FullName = stri_trans_general(FullName, "Latin-ASCII")) %>%
  rename(code = code, FullName = FullName)

code_to_name <- setNames(personnel$FullName, personnel$code)

admin <- admin %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

##########################

colz <- names(archInterp)[6:9]

code_to_name <- setNames(DS_ArchInterpPrim$description, DS_ArchInterpPrim$code)

archInterp <- archInterp %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

##########################

colz <- names(archInterp)[10:13]

code_to_name <- setNames(DS_ArchInterpAltern$description, DS_ArchInterpAltern$code)

archInterp <- archInterp %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

##########################

colz <- names(archInterp)[14:17]

code_to_name <- setNames(DS_FuncInterpPrim$description, DS_FuncInterpPrim$code)

archInterp <- archInterp %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

##########################

colz <- names(archInterp)[18:21]

code_to_name <- setNames(DS_FuncInterpAltern$description, DS_FuncInterpAltern$code)

archInterp <- archInterp %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

########################










6:21



#%>%
  select(-all_of(worker_columns))


admin_names <- admin %>%
  mutate(across(all_of(worker_columns), ~ personnel$FullName[match(., personnel$code)], .names = "name_{col}"))

  
materials_wide <- materials %>% 
  left_join(materialsLabel, by = "code") %>% 
  select(-code) %>% 
  pivot_wider(id_cols = "ID", names_from = "observation", values_from = "label")




write_csv(ceramics, paste0(DIR, "ceramics.csv"))

write_csv(lithics, paste0(DIR, "lithics.csv"))

write_csv(personnel, paste0(DIR, "personnel.csv"))


```

personnel










### Step 6: Check and reorganize totals columns (if necessary)???
This placeholder is where you'd perform checks or reordering logic

THIS WAS SKIPPED!














