---
title: "DF9 Wide"
author: "Rudolf Cesaretti"
date: "2025-01-21"
output: html_document
---

Double-check codes tables for coded variables!! If Annie was wrong in her metadata, write a paragraph about it here.

Double-check NA Values for all variables

Fix the stone quantity variable in DF9b using the data from DF8

Make a checklist for variables -- problems and change/transformations needed. Add these as columns to the metadata. Color code rows based on their status (e.g. you can basically make all of the count data green)

Remove spaces and manually edit the site and subsite variables in DF9b

Add variable for L1, L2, L3 etc site names indicating that the collection unit is not primarily Classic Period

Hierarchical classification schema for architectural and functional interps!


Nominal vs Ordinal fields
--Ordinal fields text as "1. X", "2. Y", "3. Z" etc. AND conversion to an ordered factor variable
--Nominal fields text as "X", "Y", "Z" etc. AND conversion to an unordered factor variable

make processing script generalized

convert entire processing script to a function

Process DF8,DF9,DF10,DF9b

Cross-check values!!

Then, you need to connect the dataframe to the (millon space) collection unit polygons

Using the the final corrected geospatial dataset (millon space), see if you can predict architectural and functional classifications using ML methods 
Also perform a pattern recognition analysis in the dataset!!





read_csv(paste0(DIR, "data/raw/DF9b_rev_metadata_to_assemble.csv"))
DF9b_rev_metadata_to_assemble.csv
read_csv(DF8_wide, paste0(DIR, "data/raw/DF8_wide.csv"))


# Create DF9 Wide for Cross-Checking the Data


```{r}
rm(list = ls())

```

# Load Packages

```{r}
# Load the required package
library(tidyverse)
library(readxl)
library(stringi)
library(arrow)

```


# Set Working Directory

```{r}
DIR = "C:/Users/TJ McMote/ASU Dropbox/Rudolf Cesaretti/GitHubRepos/TMP-database/"
```



### Step 0: Import and organize all excel files from the specified directory

```{r}

DF9_metadata <- read_csv(paste0(DIR, "data/raw/Metadata.csv"), locale = locale(encoding = "Latin1"))

file_list <- list.files(
  path = paste0(DIR, "data/raw/DF9_rev"),
  pattern = "\\.csv$", full.names = TRUE
)
file_names <- tools::file_path_sans_ext(basename(file_list))
df_list <- map(file_list, read_csv, locale = locale(encoding = "Latin1"))
names(df_list) <- file_names
#list2env(data_list, envir = .GlobalEnv)

data_list  <- df_list[grepl("^Data_", names(df_list))]
codes_list <- df_list[grepl("^Codes_", names(df_list))]
```

### Step 1: Merge and Clean-Up Data

```{r}

data <- data_list %>%
  discard(names(.) %in% c("Data_fieldWorkers", "Data_labAnalysts", "Data_fieldWorkers_bool_wide", "Data_labWorkers_bool_wide", "Data_Plazas")) %>%
  reduce(full_join, by = "SSN") %>% 
  mutate(labAnalyst1 = NA,
         labAnalyst2 = NA,
         labAnalyst3 = NA,
         fieldWorker1 = NA,
         fieldWorker2 = NA,
         fieldWorker3 = NA,
         fieldWorker4 = NA,
         fieldWorker5 = NA
         ) %>% 
  select(-unworkedBone) %>% 
  rename(drainSitu = drains.x,
         wallFixSitu = wallFixtures.x,
         drainSurf = drains.y,
         wallFixSurf = wallFixtures.y,
         jagueyArch = jagueys,
         jagueyMod = jaguey) %>% 
  select(DF9_metadata$varName_DF9b[!is.na(DF9_metadata$varName_DF9b)])

#DF8_wide_raw

#DF9_metadata$varName_DF9b[!is.na(DF9_metadata$varName_DF9b)]
```


### Step 2: Replace stoneCut Values with those of DF8

```{r}

DF8_stoneCut <- read_csv(paste0(DIR, "data/raw/DF8_wide_raw.csv")) %>% 
  select(SSN, CUTSTONE) %>%
  rename(stoneCut = CUTSTONE)

data <- data %>%
  select(-stoneCut) %>%
  left_join(DF8_stoneCut, by = "SSN") %>%
  select(DF9_metadata$varName_DF9b[!is.na(DF9_metadata$varName_DF9b)])

```


### Step 2.5: Stopgap Export/Import of Raw Data

```{r}
write.csv(data, paste0(DIR, "data/raw/DF9b_raw_data_v2.csv"), row.names = F, fileEncoding = "Latin1")
```

```{r}
data <- read_csv(paste0(DIR, "data/raw/DF9b_raw_data_v2.csv"), locale = locale(encoding = "Latin1")) 
```



### Step 3: Replace Coded Values with NA

```{r}

na_replacements <- DF9_metadata %>%
  filter(!is.na(codedNAValue)) %>%
  select(varName_DF9b, codedNAValue)


convert_if_numeric <- function(x) {
  temp <- suppressWarnings(as.numeric(x))
  if (all(is.na(temp) == is.na(x))) {
    return(temp)
  } else {
    return(x)
  }
}


# Convert specified values to NA
for (i in seq_len(nrow(na_replacements))) {
  col <- na_replacements$varName_DF9b[i]
  na_val <- convert_if_numeric(na_replacements$codedNAValue[i])
  
  data[[col]][data[[col]] == na_val] <- NA
}
```

### Step 4: Replace Coded Values with Descriptions

```{r}
# Recode numeric coded variables to factors using corresponding Codes_ tables.
# Assumes the metadata has a column "CodesTable" that gives the filename
# (e.g., "Codes_A.csv") for variables that need recoding.

coded_vars <- DF9_metadata %>% 
  filter(!is.na(codesTable_DF9b) & codesTable_DF9b != "")

for(i in seq_len(nrow(coded_vars))) {
  var_name <- coded_vars$varName_DF9b[i]
  codes_table <- coded_vars$codesTable_DF9b[i]
  
  # Read in the appropriate codes mapping.
  # Assumes each codes file has at least two columns: 
  # one with the numeric code (named "Code") and one with the description (named "Description").
  codes_mapping <- codes_list[[codes_table]]
  
  # Recode the variable:
  # Convert the raw numeric codes to factors with levels defined by the "Code" column
  # and labels from the "Description" column.
  data <- data %>% 
    mutate(!!sym(var_name) := factor(!!sym(var_name),
                                      levels = codes_mapping$code,
                                      labels = codes_mapping$description))
}

data_fct <- data

data_char <- data %>% mutate(across(where(is.factor), as.character))
```


### Step 5: Revert Personnel Variables to original DF9 Form

```{r}

DF8_wide <- read_csv(paste0(DIR, "data/raw/DF8_wide_raw.csv"))

DF8_wide %>% select(SSN, FLDWRK1, FLDWRK2, FLDWRK3, FLDWRK4, FLDWRK5, CERANWK1, CERANWK2, CERANWK3) %>% 
  rename(fieldWorker1 = FLDWRK1,
         fieldWorker2 = FLDWRK2,
         fieldWorker3 = FLDWRK3,
         fieldWorker4 = FLDWRK4,
         fieldWorker5 = FLDWRK5,
         ceramicAnalyst1 = CERANWK1,
         ceramicAnalyst2 = CERANWK2,
         ceramicAnalyst3 = CERANWK3) %>% 
  mutate(across(everything(), ~ ifelse(. == -1, NA, .)))-> personnel

recode_vector <- setNames(codes_list[["Codes_personnel"]]$description, codes_list[["Codes_personnel"]]$code)

recode_column <- function(column, recode_vector) {
  factor(column, levels = as.numeric(names(recode_vector)), labels = recode_vector[as.numeric(names(recode_vector))])
}

personnel <- personnel %>%
  mutate(
    across(
      c(fieldWorker1, fieldWorker2, fieldWorker3, fieldWorker4, fieldWorker5,
        ceramicAnalyst1, ceramicAnalyst2, ceramicAnalyst3),
      recode_column, recode_vector = recode_vector
    )
  )

personnel_char <- personnel %>% mutate(across(where(is.factor), as.character)) %>% 
  rename(labAnalyst1 = ceramicAnalyst1, 
         labAnalyst2 = ceramicAnalyst2, 
         labAnalyst3 = ceramicAnalyst3)

write.csv(personnel_char, paste0(DIR, "data/raw/personnel.csv"), fileEncoding = "Latin1")

data_char <- data_char %>%
  select(-c(fieldWorker1, fieldWorker2, fieldWorker3, fieldWorker4, fieldWorker5,
            labAnalyst1, labAnalyst2, labAnalyst3)) %>% 
  left_join(personnel_char, by = "SSN") %>%
  select(DF9_metadata$varName_DF9b[!is.na(DF9_metadata$varName_DF9b)])

write.csv(data_char, paste0(DIR, "data/raw/data_char.csv"), row.names = F, fileEncoding = "Latin1")
```


### Step 6: Export Data

```{r}
write.csv(data_char, paste0(DIR, "data/raw/DF9b_data_char.csv"), row.names = F, fileEncoding = "Latin1")
```


```{r}
write.csv(data_char, paste0(DIR, "data/raw/DF9b_data_char.csv"), row.names = F, fileEncoding = "Latin1")
```





# UNUSED DRAFTINGS




FIRST, CHANGE ORDINAL FACTOR VARIABLE DESCRIPTIONS SO THAT THEY INCLUDE THE NUMBER IN FRONT OF THE DESCRIPTION

We can extend this workflow by incorporating a metadata column (say, `FactorType`) that indicates whether a coded variable should be treated as ordinal (ordered) or nominal (unordered). Then, when you convert the numeric codes to factors, you simply check that metadata value and set the `ordered` argument in the `factor()` function accordingly.

Below is an example that builds on the previous recoding loop. In this snippet, the metadata is assumed to have:

*   A column (e.g., `DF9b_VarName`) with the variable name,
*   A column `CodesTable` indicating which "Codes\_" table to use,
*   And a column `FactorType` which holds values like `"ordered"` or `"unordered"`.

```{r}

# Assume DF9_metadata contains columns:
# DF9b_VarName, CodesTable, and FactorType.

# Filter to only those variables that need recoding.
coded_vars <- DF9_metadata %>% 
  filter(!is.na(CodesTable) & CodesTable != "")

for(i in seq_len(nrow(coded_vars))) {
  var_name   <- coded_vars$DF9b_VarName[i]
  codes_file <- coded_vars$CodesTable[i]
  factor_type <- coded_vars$FactorType[i]  # Expected to be "ordered" or "unordered"
  
  # Read the corresponding codes mapping
  codes_mapping <- read_csv(codes_file) %>% 
    select(Code, Description)
  
  # Determine if the factor should be ordered
  is_ordered <- tolower(factor_type) == "ordered"
  
  # Recode the variable using the codes mapping and set factor order accordingly.
  data <- data %>% 
    mutate(!!sym(var_name) := factor(!!sym(var_name),
                                     levels = codes_mapping$Code,
                                     labels = codes_mapping$Description,
                                     ordered = is_ordered))
}
```



```{r}
str(data)
```


```{r}

```

























# EDA and Data Summarization


Reasoned about dataset summarization tools for 19 seconds

Below are eight separate R scripts—one per package—to load your CSV file and generate a data summary or metadata report. In each script, we assume your CSV file is named `"DF9b_raw_data.csv"`. Adjust file paths or function options as needed.

* * *

### Script 1: Using **summarytools**

```{r}
# Script 1: Dataset Summarization using summarytools
library(summarytools)

# Read the dataset
#data <- read_csv("DF9b_raw_data.csv")

# Generate a detailed summary report of the dataset
summary_report <- dfSummary(
  data,
  plain.ascii = FALSE, 
  style = "grid",
  valid.col = FALSE,
  headings = TRUE
)

# Print the summary to the console
print(summary_report)

# Optionally, view the report in a browser/HTML viewer
view(summary_report)
```

* * *

### Script 2: Using **dlookr**

```{r}
# Script 2: Dataset Diagnosis using dlookr
library(dlookr)

# Read the dataset
# Generate a diagnostic report (includes missing values, distributions, outliers, etc.)
# The report is exported as an HTML file.
diagnose_report(
  data, 
  output_format = "html", 
  output_file = "dlookr_report.html"
)
```

* * *

### Script 3: Using **codebook**

```{r}
# Script 3: Generating Codebook using the codebook package
library(codebook)

# Read the dataset
# Generate and print a codebook for the dataset
# This function automatically creates a codebook with variable descriptions and summary statistics.
codebook(data)
```

### Script 4: Using **skimr**

```{r}
# Script 4: Quick Data Summary using skimr
library(skimr)


# Obtain a comprehensive summary of the dataset (including missing values, distribution, etc.)
skim(data)
```


### Script 5: Using **dataMeta**

> _Note:_ The exact function name can vary with version. Here we assume a function called `create_meta()` exists for extracting metadata.

```{r}
# Script 5: Extracting Metadata using dataMeta
library(tableone)


# Extract metadata information from the dataset
# (Check your version’s documentation if the function name differs.)
tableOne <- CreateTableOne(data=data)

tableOne
print(tableOne)
summary(tableOne)

## See the categorical part only using $ operator
tableOne$CatTable
summary(tableOne$CatTable)

## See the continuous part only using $ operator
tableOne$ContTable
summary(tableOne$ContTable)

```

* * *

### Script 6: Using **DataExplorer**

```{r}
# Script 6: Exploratory Data Analysis Report using DataExplorer
library(DataExplorer)

# Create an automated EDA report in HTML format
create_report(data, output_file = "DataExplorer_report.html")
```

* * *

### Script 7: Using **explore**

```{r}
# Script 7: Interactive Data Exploration using explore package
library(explore)

# Launch an interactive exploration of the dataset
# This typically opens an interactive report in your RStudio viewer or default browser.
explore(data)
```

* * *

### Script 8: Using **SmartEDA**

```{r}
# Script 8: Automated Exploratory Data Analysis with SmartEDA
library(SmartEDA)

# Generate an EDA report (HTML) using SmartEDA
ExpReport(data, op_file = "SmartEDA_report.html", op_dir = getwd())
```

* * *

Each of these scripts loads the data and then leverages a different package to summarize your dataset or extract metadata. You can run these individually or combine parts of them into your workflow as needed.



"admin" is the base table

-- 5050-row tables --

WorkingHierarchy
`a_FuncInterp1 for digitization`
archInterp
archMaterial




```{r}

ceramics = cerPhTot %>% 
  left_join(cerVessel, by = "ID") %>% 
  left_join(cerNonVessel, by = "ID")

lithics = lithicFlaked %>% 
  left_join(lithicGround, by = "ID")

artifactOther

```






complexData
complexMacroData

condition
description

location

Plazas
workshop



materials
materialsLabel


z_archInterp1
z_archInterp2
z_funcInterp1
z_funcInterp2
z_vegetation
HierarchyCode



personnelLabel
"personnel error check"
fieldWorkers
labAnalysts


```{r}




colz <- c("fieldWorker1", "fieldWorker2", "fieldWorker3", "fieldWorker4", "fieldWorker5",
                    "ceramicAnalyst1", "ceramicAnalyst2", "ceramicAnalyst3")


personnel <- personnelLabel %>% 
  mutate(FullName = paste(first_name, last_name, sep = " ")) %>% 
  select(-last_name, -first_name) %>% 
  mutate(FullName = stri_trans_general(FullName, "Latin-ASCII")) %>%
  rename(code = code, FullName = FullName)

code_to_name <- setNames(personnel$FullName, personnel$code)

admin <- admin %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

##########################

colz <- names(archInterp)[6:9]

code_to_name <- setNames(DS_ArchInterpPrim$description, DS_ArchInterpPrim$code)

archInterp <- archInterp %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

##########################

colz <- names(archInterp)[10:13]

code_to_name <- setNames(DS_ArchInterpAltern$description, DS_ArchInterpAltern$code)

archInterp <- archInterp %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

##########################

colz <- names(archInterp)[14:17]

code_to_name <- setNames(DS_FuncInterpPrim$description, DS_FuncInterpPrim$code)

archInterp <- archInterp %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

##########################

colz <- names(archInterp)[18:21]

code_to_name <- setNames(DS_FuncInterpAltern$description, DS_FuncInterpAltern$code)

archInterp <- archInterp %>%
  mutate(across(all_of(colz), ~ code_to_name[as.character(.)])) 

########################










6:21



#%>%
  select(-all_of(worker_columns))


admin_names <- admin %>%
  mutate(across(all_of(worker_columns), ~ personnel$FullName[match(., personnel$code)], .names = "name_{col}"))

  
materials_wide <- materials %>% 
  left_join(materialsLabel, by = "code") %>% 
  select(-code) %>% 
  pivot_wider(id_cols = "ID", names_from = "observation", values_from = "label")




write_csv(ceramics, paste0(DIR, "ceramics.csv"))

write_csv(lithics, paste0(DIR, "lithics.csv"))

write_csv(personnel, paste0(DIR, "personnel.csv"))


```

personnel










### Step 6: Check and reorganize totals columns (if necessary)???
This placeholder is where you'd perform checks or reordering logic

THIS WAS SKIPPED!














