---
title: "EDA and Data Summarization"
author: "Rudolf Cesaretti"
date: "2025-03-07"
output: html_document
---


outlier

pattern


exploreR
funModeling
xray

https://www.analyticsvidhya.com/blog/2022/10/three-r-libraries-for-automated-eda/
https://journal.r-project.org/articles/RJ-2019-033/

## Setup & Data 

```{r}
rm(list = ls())
```


```{r}
library(tidyverse)
library(readxl)
library(stringi)
```

```{r}
DIR = "C:/Users/TJ McMote/ASU Dropbox/Rudolf Cesaretti/GitHubRepos/TMP-database/"

```


```{r}
metadata <- read_csv(paste0(DIR, "data/raw/Metadata.csv"), locale = locale(encoding = "Latin1"))


file_list <- list.files(
  path = paste0(DIR, "data/raw/DF9_rev"),
  pattern = "^Data_.*\\.csv$", 
  full.names = TRUE
)
file_names <- tools::file_path_sans_ext(basename(file_list))
data_list <- map(file_list, read_csv, locale = locale(encoding = "Latin1"))
names(data_list) <- file_names
#list2env(data_list, envir = .GlobalEnv)

data <- read_csv(paste0(DIR, "data/raw/DF9b_data_char.csv"), locale = locale(encoding = "Latin1"))
```


```{r}

dataReportData_list <- list()

NOM <- c("complexData", "complexMacroData", "archInterp", "archMaterial", "condition", "description", "workshop", "admin", "location","lithicGround","lithicFlaked", "cerVessel", "cerPhTot", "cerNonVessel", "artifactOther")

for (i in NOM) {
  #nom2 <- paste0("Data_", i)
  dataReportData_list[[i]] <- data %>% 
    select(metadata %>% filter(dbTable_DF9b == paste0("Data_", i)) %>% select(varName_DF9b) %>% pull())
  
}

dataReportData_list[["Plazas"]] <- data_list[["Data_Plazas"]] %>% 
  select(-SSN) %>% 
  mutate(NorwoodPlazas = ifelse(PlazaNo < 0, FALSE, TRUE),
         PlazaNo = ifelse(PlazaNo == -1, NA, PlazaNo)) %>% 
  rename(PlazaNum = PlazaNo)

SSN <- data.frame(SSN = data$SSN)

dataReportData_list[["labWorkers"]] <- SSN %>% 
  left_join(data_list[["Data_labWorkers_bool_wide"]], by = "SSN") %>% 
  select(-SSN)

dataReportData_list[["fieldWorkers"]] <- SSN %>% 
  left_join(data_list[["Data_fieldWorkers_bool_wide"]], by = "SSN") %>% 
  select(-SSN)

```

```{r}
data_list <- dataReportData_list

rm(dataReportData_list, file_names, file_list, NOM, i, SSN)
```






## EDA and Data Summarization


Reasoned about dataset summarization tools for 19 seconds

Below are eight separate R scripts—one per package—to load your CSV file and generate a data summary or metadata report. In each script, we assume your CSV file is named `"DF9b_raw_data.csv"`. Adjust file paths or function options as needed.

* * *

### summarytools

summarytools: Tools to Quickly and Neatly Summarize Data

Data frame summaries, cross-tabulations, weight-enabled frequency tables and common descriptive (univariate) statistics in concise tables available in a variety of formats (plain ASCII, Markdown and HTML). A good point-of-entry for exploring data, both for experienced and new R users.

descr and dfSummary, from the summarytools package

https://github.com/dcomtois/summarytools
https://cran.r-project.org/web/packages/summarytools/index.html
https://cran.r-project.org/web/packages/summarytools/vignettes/introduction.html
https://cran.r-project.org/web/packages/summarytools/vignettes/rmarkdown.html
https://cran.r-project.org/web/packages/summarytools/summarytools.pdf
https://dcomtois.r-universe.dev/summarytools

```{r}
# Script 1: Dataset Summarization using summarytools
library(summarytools)

# dataset
DAT <- data_list[[1]]

# Generate a detailed summary report of the dataset
summary_report <- dfSummary(
  DAT,
  plain.ascii = FALSE, 
  style = "multiline",
  valid.col = FALSE,
  headings = TRUE,
  trim.strings = FALSE,
  max.distinct.values = 50,
  max.string.width = 50,
)
summary_report
# Print the summary to the console
print(summary_report)

# Optionally, view the report in a browser/HTML viewer
view(summary_report)
```

* * *

### dlookr

dlookr: Tools for Data Diagnosis, Exploration, Transformation

A collection of tools that support data diagnosis, exploration, and transformation. Data diagnostics provides information and visualization of missing values, outliers, and unique and negative values to help you understand the distribution and quality of your data. Data exploration provides information and visualization of the descriptive statistics of univariate variables, normality tests and outliers, correlation of two variables, and the relationship between the target variable and predictor. Data transformation supports binning for categorizing continuous variables, imputes missing values and outliers, and resolves skewness. And it creates automated reports that support these three tasks.

https://choonghyunryu.github.io/dlookr/
https://choonghyunryu.r-universe.dev/dlookr
https://github.com/choonghyunryu/dlookr
https://cran.r-project.org/web/packages/dlookr/index.html
https://cran.r-project.org/web/packages/dlookr/dlookr.pdf
https://cran.r-project.org/web/packages/dlookr/vignettes/introduce.html
https://cran.r-project.org/web/packages/dlookr/vignettes/EDA.html
https://cran.r-project.org/web/packages/dlookr/vignettes/diagonosis.html
https://cran.r-project.org/web/packages/dlookr/vignettes/transformation.html


```{r}
# Script 2: Dataset Diagnosis using dlookr
library(dlookr)


# dataset
DAT <- data_list[[1]]
# Read the dataset
# Generate a diagnostic report (includes missing values, distributions, outliers, etc.)
# The report is exported as an HTML file.
diagnose_report(
  data, 
  output_format = "html", 
  output_file = "dlookr_report.html"
)

diagnose_category(DAT)
diagnose_outlier()
plot_outlier()
```

* * *

### Using codebook

codebook: Automatic Codebooks from Metadata Encoded in Dataset Attributes

Easily automate the following tasks to describe data frames: Summarise the distributions, and labelled missings of variables graphically and using descriptive statistics. For surveys, compute and summarise reliabilities (internal consistencies, retest, multilevel) for psychological scales. Combine this information with metadata (such as item labels and labelled values) that is derived from R attributes. To do so, the package relies on 'rmarkdown' partials, so you can generate HTML, PDF, and Word documents. Codebooks are also available as tables (CSV, Excel, etc.) and in JSON-LD, so that search engines can find your data and index the metadata. The metadata are also available at your fingertips via RStudio Addins.

https://rubenarslan.github.io/codebook/
https://cloud.r-project.org/web/packages/codebook/index.html
https://cloud.r-project.org/web/packages/codebook/codebook.pdf
https://cloud.r-project.org/web/packages/codebook/vignettes/codebook_tutorial.html
https://cloud.r-project.org/web/packages/codebook/vignettes/codebook.html
https://cloud.r-project.org/web/packages/codebook/vignettes/codebook_qualtrics.html
https://cloud.r-project.org/web/packages/codebook/vignettes/codebook_sav.html
https://github.com/rubenarslan/codebook


```{r}
# Script 3: Generating Codebook using the codebook package
library(codebook)

# Read the dataset
# Generate and print a codebook for the dataset
# This function automatically creates a codebook with variable descriptions and summary statistics.
codebook(data)
```

### skimr

skimr: Compact and Flexible Summaries of Data

A simple to use summary function that can be used with pipes and displays nicely in the console. The default summary statistics may be modified by the user as can the default formatting. Support for data frames and vectors is included, and users can implement their own skim methods for specific object types as described in a vignette. Default summaries include support for inline spark graphs. Instructions for managing these on specific operating systems are given in the "Using skimr" vignette and the README.

https://docs.ropensci.org/skimr/
https://cloud.r-project.org/web/packages/skimr/index.html
https://cloud.r-project.org/web/packages/skimr/vignettes/Skimr_defaults.html
https://cloud.r-project.org/web/packages/skimr/vignettes/Using_fonts.html
https://cloud.r-project.org/web/packages/skimr/vignettes/extending_skimr.html
https://cloud.r-project.org/web/packages/skimr/vignettes/skimr.html
https://cloud.r-project.org/web/packages/skimr/skimr.pdf
https://github.com/ropensci/skimr/

```{r}
# Script 4: Quick Data Summary using skimr
library(skimr)


# Obtain a comprehensive summary of the dataset (including missing values, distribution, etc.)
skim(data)
```


### dataMeta

dataMeta: Create and Append a Data Dictionary for an R Dataset

Designed to create a basic data dictionary and append to the original dataset's attributes list. The package makes use of a tidy dataset and creates a data frame that will serve as a linker that will aid in building the dictionary. The dictionary is then appended to the list of the original dataset's attributes. The user will have the option of entering variable and item descriptions by writing code or use alternate functions that will prompt the user to add these.

https://github.com/dmrodz/dataMeta
https://cran.r-project.org/web/packages/dataMeta/index.html
https://cran.r-project.org/web/packages/dataMeta/vignettes/dataMeta_Vignette.html
https://cran.r-project.org/web/packages/dataMeta/dataMeta.pdf
https://dmrodz.r-universe.dev/dataMeta

> _Note:_ The exact function name can vary with version. Here we assume a function called `create_meta()` exists for extracting metadata.

```{r}

```

### tableone

tableone: Create 'Table 1' to Describe Baseline Characteristics with or without Propensity Score Weights

Creates 'Table 1', i.e., description of baseline patient characteristics, which is essential in every medical research. Supports both continuous and categorical variables, as well as p-values and standardized mean differences. Weighted data are supported via the 'survey' package.

https://kaz-yos.r-universe.dev/tableone
https://github.com/kaz-yos/tableone
https://cran.r-project.org/web/packages/tableone/index.html
https://cran.r-project.org/web/packages/tableone/vignettes/introduction.html
https://cran.r-project.org/web/packages/tableone/vignettes/smd.html
https://cran.r-project.org/web/packages/tableone/tableone.pdf

```{r}
# Script 5: Extracting Metadata using dataMeta
library(tableone)


# Extract metadata information from the dataset
# (Check your version’s documentation if the function name differs.)
tableOne <- CreateTableOne(data=data, includeNA=TRUE, addOverall=TRUE)

tableOne
print(tableOne, showAllLevels = TRUE, formatOptions = list(big.mark = ","))
summary(tableOne)



## See the categorical part only using $ operator
tableOne$CatTable
summary(tableOne$CatTable)

## See the continuous part only using $ operator
tableOne$ContTable
summary(tableOne$ContTable)

```

* * *

### DataExplorer

DataExplorer: Automate Data Exploration and Treatment

Automated data exploration process for analytic tasks and predictive modeling, so that users could focus on understanding data and extracting insights. The package scans and analyzes each variable, and visualizes them with typical graphical techniques. Common data processing methods are also available to treat and format data.

http://boxuancui.github.io/DataExplorer/
https://cran.r-project.org/package=DataExplorer
https://github.com/boxuancui/DataExplorer/
https://cloud.r-project.org/web/packages/DataExplorer/index.html
https://cloud.r-project.org/web/packages/DataExplorer/DataExplorer.pdf
https://cloud.r-project.org/web/packages/DataExplorer/vignettes/dataexplorer-intro.html

```{r}
# Script 6: Exploratory Data Analysis Report using DataExplorer
library(DataExplorer)

# Create an automated EDA report in HTML format
create_report(dataReportData_list[[1]], output_file = "DataExplorer_report.html")
```

* * *

### explore

explore: Simplifies Exploratory Data Analysis

Interactive data exploration with one line of code, automated reporting or use an easy to remember set of tidy functions for low code exploratory data analysis.

Simplifies Exploratory Data Analysis:

Interactive data exploration: explore()
Use AI to unveil hidden patterns in your data (xgboost, RF, logreg, DT): explain_*()
Generate an automated report of your data (or patterns in your data): report()
Manual exploration: explore(), describe(), explain_*(), abtest(), …
18 ready to use datasets for teaching & testing: use_data_*(), create_data_*()

https://rolkra.github.io/explore/
https://rolkra.r-universe.dev/explore
https://cran.r-project.org/web/packages/explore/index.html
https://cran.r-project.org/web/packages/explore/explore.pdf
https://cran.r-project.org/web/packages/explore/vignettes/explore.html
https://cran.r-project.org/web/packages/explore/vignettes/abtest.html
https://cran.r-project.org/web/packages/explore/vignettes/clean-drop.html
https://cran.r-project.org/web/packages/explore/vignettes/data.html
https://cran.r-project.org/web/packages/explore/vignettes/describe.html
https://cran.r-project.org/web/packages/explore/vignettes/explain.html
https://cran.r-project.org/web/packages/explore/vignettes/explore-mtcars.html
https://cran.r-project.org/web/packages/explore/vignettes/explore-penguins.html
https://cran.r-project.org/web/packages/explore/vignettes/explore-titanic.html
https://cran.r-project.org/web/packages/explore/vignettes/predict.html
https://cran.r-project.org/web/packages/explore/vignettes/report-target.html
https://cran.r-project.org/web/packages/explore/vignettes/report-targetpct.html
https://cran.r-project.org/web/packages/explore/vignettes/report.html
https://cran.r-project.org/web/packages/explore/vignettes/tips-tricks.html

```{r}
# Script 7: Interactive Data Exploration using explore package
library(explore)

# Launch an interactive exploration of the dataset
# This typically opens an interactive report in your RStudio viewer or default browser.
explore(data)
```

* * *

### SmartEDA

SmartEDA: Summarize and Explore the Data

Exploratory analysis on any input data describing the structure and the relationships present in the data. The package automatically select the variable and does related descriptive statistics. Analyzing information value, weight of evidence, custom tables, summary statistics, graphical techniques will be performed for both numeric and categorical predictors.

https://daya6489.github.io/SmartEDA/
https://cloud.r-project.org/web/packages/SmartEDA/vignettes/SmartEDA.html
https://cloud.r-project.org/web/packages/SmartEDA/vignettes/SmartTwoPlots.html
https://cloud.r-project.org/web/packages/SmartEDA/vignettes/CustomTable.html
https://cloud.r-project.org/web/packages/SmartEDA/SmartEDA.pdf


```{r}
# Script 8: Automated Exploratory Data Analysis with SmartEDA
library(SmartEDA)

# Generate an EDA report (HTML) using SmartEDA
ExpReport(
  data = data,
  Template = NULL, #R markdown template (.rmd file)
  Target = NULL, #dependent variable. If there is no defined target variable then keep as it is NULL.
  label = NULL, #target variable descriptions, not a mandatory field
  theme = "Default",
  op_file = "SmartEDA_report.html",
  op_dir = paste0(DIR, "reports/"),
  sc = NULL, #sample number of plots for categorical variable. User can decide how many number of plots to depict in html report.
  sn = NULL, #sample number of plots for numerical variable. User can decide how many number of plots to depict in html report.
  Rc = NULL   #reference category of target variable. If Target is categorical then Pclass value is mandatory and which should not be NULL
)


```

* * *

### ExPanDaR

ExPanDaR: Explore Your Data Interactively

Provides a shiny-based front end (the 'ExPanD' app) and a set of functions for exploratory data analysis. Run as a web-based app, 'ExPanD' enables users to assess the robustness of empirical evidence without providing them access to the underlying data. You can export a notebook containing the analysis of 'ExPanD' and/or use the functions of the package to support your exploratory data analysis workflow. Refer to the vignettes of the package for more information on how to use 'ExPanD' and/or the functions of this package.

https://joachim-gassen.github.io/ExPanDaR/
https://cran.r-project.org/package=ExPanDaR
https://joachim-gassen.r-universe.dev/ExPanDaR

```{r}
library(ExPanDaR)
```



### naniar

naniar: Data Structures, Summaries, and Visualisations for Missing Data

Missing values are ubiquitous in data and need to be explored and handled in the initial stages of analysis. 'naniar' provides data structures and functions that facilitate the plotting of missing values and examination of imputations. This allows missing data dependencies to be explored with minimal deviation from the common work patterns of 'ggplot2' and tidy data. The work is fully discussed at Tierney & Cook (2023) <doi:10.18637/jss.v105.i07>.

https://njtierney.r-universe.dev/naniar
https://naniar.njtierney.com/
https://github.com/njtierney/naniar
https://cran.r-project.org/package=naniar

```{r}
library(naniar)
```


### dataMaid

dataMaid: A Suite of Checks for Identification of Potential Errors in a Data Frame as Part of the Data Screening Process

Data screening is an important first step of any statistical analysis. dataMaid auto generates a customizable data report with a thorough summary of the checks and the results that a human can use to identify possible errors. It provides an extendable suite of test for common potential errors in a dataset.

https://github.com/ekstroem/dataMaid
https://cran.r-project.org/web/packages/dataMaid/index.html
https://cran.r-project.org/web/packages/dataMaid/dataMaid.pdf
https://cran.r-project.org/web/packages/dataMaid/vignettes/extending_dataMaid.html
https://sandsynligvis.dk/2017/08/21/datamaid-your-personal-assistant-for-cleaning-up-the-data-cleaning-process/
https://sandsynligvis.dk/2018/03/03/generating-codebooks-in-r/
https://www.jstatsoft.org/article/view/v090i06

```{r}
library(dataMaid)
```


### dataReporter

dataReporter: Reproducible Data Screening Checks and Report of Possible Errors

Data screening is an important first step of any statistical analysis. 'dataReporter' auto generates a customizable data report with a thorough summary of the checks and the results that a human can use to identify possible errors. It provides an extendable suite of test for common potential errors in a dataset. See Petersen AH, Ekstrøm CT (2019). "dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R." _Journal of Statistical Software_, *90*(6), 1-38 <doi:10.18637/jss.v090.i06> for more information.

https://github.com/ekstroem/dataReporter
https://cran.r-project.org/web/packages/dataReporter/index.html
https://cran.r-project.org/web/packages/dataReporter/dataReporter.pdf
https://cran.r-project.org/web/packages/dataReporter/vignettes/extending_dataReporter.html
https://rbasics.org/packages/datareporter-package-in-r/
https://ekstroem.r-universe.dev/dataReporter

```{r}
library(dataReporter)
```


### correlationfunnel

correlationfunnel: Speed Up Exploratory Data Analysis (EDA) with the Correlation Funnel
Speeds up exploratory data analysis (EDA) by providing a succinct workflow and interactive visualization tools for understanding which features have relationships to target (response). Uses binary correlation analysis to determine relationship. Default correlation method is the Pearson method. Lian Duan, W Nick Street, Yanchi Liu, Songhua Xu, and Brook Wu (2014) <doi:10.1145/2637484>.

https://github.com/business-science/correlationfunnel
https://cran.r-project.org/web/packages/correlationfunnel/
https://cran.r-project.org/web/packages/correlationfunnel/vignettes/introducing_correlation_funnel.html
https://cran.r-project.org/web/packages/correlationfunnel/vignettes/key_considerations.html
https://cran.r-project.org/web/packages/correlationfunnel/correlationfunnel.pdf

```{r}
library(correlationfunnel)
```


### janitor

janitor: Simple Tools for Examining and Cleaning Dirty Data

The main janitor functions can: perfectly format data.frame column names; provide quick counts of variable combinations (i.e., frequency tables and crosstabs); and explore duplicate records. Other janitor functions nicely format the tabulation results. These tabulate-and-report functions approximate popular features of SPSS and Microsoft Excel. This package follows the principles of the "tidyverse" and works well with the pipe function %>%. janitor was built with beginning-to-intermediate R users in mind and is optimized for user-friendliness.

https://sfirke.github.io/janitor/
https://sfirke.r-universe.dev/janitor
https://github.com/sfirke/janitor/
https://cran.r-project.org/package=janitor
https://cloud.r-project.org/web/packages/janitor/janitor.pdf
https://cran.r-project.org/web/packages/janitor/vignettes/janitor.html
https://cran.r-project.org/web/packages/janitor/vignettes/tabyls.html

```{r}
library(janitor)

```



### data.validator

data.validator: Automatic Data Validation and Reporting

Validate dataset by columns and rows using convenient predicates inspired by 'assertr' package. Generate good looking HTML report or print console output to display in logs of your data processing pipeline.

https://appsilon.github.io/data.validator/
https://cran.r-project.org/package=data.validator
https://appsilon.r-universe.dev/data.validator
https://github.com/Appsilon/data.validator


```{r}
library(data.validator)
```



### pointblank

pointblank: Data Validation and Organization of Metadata for Local and Remote Tables

Validate data in data frames, 'tibble' objects, 'Spark' 'DataFrames', and database tables. Validation pipelines can be made using easily-readable, consecutive validation steps. Upon execution of the validation plan, several reporting options are available. User-defined thresholds for failure rates allow for the determination of appropriate reporting actions. Many other workflows are available including an information management workflow, where the aim is to record, collect, and generate useful information on data tables.

https://rstudio.github.io/pointblank/
https://rstudio.r-universe.dev/pointblank
https://cran.r-project.org/package=pointblank

```{r}
library(pointblank)
```


### origami


origami: Generalized Framework for Cross-Validation

A general framework for the application of cross-validation schemes to particular functions. By allowing arbitrary lists of results, origami accommodates a range of cross-validation applications. This implementation was first described by Coyle and Hejazi (2018) <doi:10.21105/joss.00512>.

https://cran.r-project.org/package=origami
https://tlverse.org/origami/
https://tlverse.r-universe.dev/origami
https://github.com/tlverse/origami

```{r}
library(origami)
```


### extremevalues


extremevalues: Univariate Outlier Detection
Detect outliers in one-dimensional data.

https://markvanderloo.r-universe.dev/extremevalues
https://cran.r-project.org/package=extremevalues
https://cran.r-project.org/web/packages/extremevalues/vignettes/SN201003.pdf
https://cran.r-project.org/web/packages/extremevalues/vignettes/SN201003.pdf
https://markvanderloo.r-universe.dev/extremevalues

```{r}
library(extremevalues)
```


### synthesizer


synthesizer: Fast, Robust, and High-Quality Synthetic Data Generation with a Tuneable Privacy-Utility Trade-Off

Synthesize numeric, categorical, mixed and time series data. Data circumstances including mixed (or zero-inflated) distributions and missing data patterns are reproduced in the synthetic data. A single parameter allows balancing between high-quality synthetic data that represents correlations of the original data and lower quality but more privacy safe synthetic data without correlations. Tuning can be done per variable or for the whole dataset.

https://markvanderloo.r-universe.dev/synthesizer
https://github.com/markvanderloo/synthesizer
https://cran.r-project.org/web/packages/synthesizer/index.html
https://cran.r-project.org/web/packages/synthesizer/vignettes/introduction.html
https://cran.r-project.org/web/packages/synthesizer/synthesizer.pdf


```{r}
library(synthesizer)
```


### lumberjack

lumberjack: Track Changes in Data

A framework that allows for easy logging of changes in data. Main features: start tracking changes by adding a single line of code to an existing script. Track changes in multiple datasets, using multiple loggers. Add custom-built loggers or use loggers offered by other packages. <doi:10.18637/jss.v098.i01>.

https://markvanderloo.r-universe.dev/lumberjack
https://cran.r-project.org/package=lumberjack
https://cran.r-project.org/web/packages/lumberjack/vignettes/using_lumberjack.pdf
https://cran.r-project.org/web/packages/lumberjack/vignettes/JSS_4008.pdf
https://cran.r-project.org/web/packages/lumberjack/lumberjack.pdf
https://github.com/markvanderloo/lumberjack

```{r}
library(lumberjack)
```


### mde

mde: Missing Data Explorer
Correct identification and handling of missing data is one of the most important steps in any analysis. To aid this process, 'mde' provides a very easy to use yet robust framework to quickly get an idea of where the missing data lies and therefore find the most appropriate action to take. Graham WJ (2009) <doi:10.1146/annurev.psych.58.110405.085530>.

https://nelson-gon.r-universe.dev/mde
https://cran.r-project.org/package=mde
https://github.com/nelson-gon/mde

```{r}
library(mde)
```


### loon

loon: Interactive Statistical Data Visualization
An extendable toolkit for interactive data visualization and exploration.

https://great-northern-diver.r-universe.dev/loon
https://great-northern-diver.github.io/loon/
https://cran.r-project.org/package=loon
https://github.com/great-northern-diver/loon

```{r}
library(loon)
```



### tldr

tldr: T Loux Doing R: Functions to Simplify Data Analysis and Reporting
Gives a number of functions to aid common data analysis processes and reporting statistical results in an 'RMarkdown' file. Data analysis functions combine multiple base R functions used to describe simple bivariate relationships into a single, easy to use function. Reporting functions will return character strings to report p-values, confidence intervals, and hypothesis test and regression results. Strings will be LaTeX-formatted as necessary and will knit pretty in an 'RMarkdown' document. The package also provides wrappers function in the 'tableone' package to make the results knit-able.

https://tloux.r-universe.dev/tldr
https://cran.r-project.org/package=tldr
https://github.com/tloux/tldr
```{r}
library(tldr)
```


### Hmisc

Hmisc: Harrell Miscellaneous
Contains many functions useful for data analysis, high-level graphics, utility operations, functions for computing sample size and power, simulation, importing and annotating datasets, imputing missing values, advanced table making, variable clustering, character string manipulation, conversion of R objects to LaTeX and html code, recoding variables, caching, simplified parallel computing, encrypting and decrypting data using a safe workflow, general moving window statistical estimation, and assistance in interpreting principal component analysis.

https://harrelfe.r-universe.dev/Hmisc
https://cran.r-project.org/package=Hmisc
https://github.com/harrelfe/Hmisc

describe, from the Hmisc package

```{r}
library(Hmisc)
```


### arsenal

arsenal: An Arsenal of 'R' Functions for Large-Scale Statistical Summaries

An Arsenal of 'R' functions for large-scale statistical summaries, which are streamlined to work within the latest reporting tools in 'R' and 'RStudio' and which use formulas and versatile summary statistics for summary tables and models. The primary functions include tableby(), a Table-1-like summary of multiple variable types 'by' the levels of one or more categorical variables; paired(), a Table-1-like summary of multiple variable types paired across two time points; modelsum(), which performs simple model fits on one or more endpoints for many variables (univariate or adjusted for covariates); freqlist(), a powerful frequency table across many categorical variables; comparedf(), a function for comparing data.frames; and write2(), a function to output tables to a document.

https://mayoverse.r-universe.dev/arsenal
https://mayoverse.github.io/arsenal/
https://github.com/mayoverse/arsenal
https://cran.r-project.org/package=arsenal


```{r}
library(arsenal)
```


### dataCompareR

dataCompareR: Compare Two Data Frames and Summarise the Difference

Easy comparison of two tabular data objects in R. Specifically designed to show differences between two sets of data in a useful way that should make it easier to understand the differences, and if necessary, help you work out how to remedy them. Aims to offer a more useful output than all.equal() when your two data sets do not match, but isn't intended to replace all.equal() as a way to test for equality.

https://capitalone.github.io/dataCompareR/index.html
https://cloud.r-project.org/web/packages/dataCompareR/index.html
https://capitalone.r-universe.dev/dataCompareR
https://github.com/capitalone/dataCompareR

```{r}
library(dataCompareR)
```


### forcats

forcats: Tools for Working with Categorical Variables (Factors)

Helpers for reordering factor levels (including moving specified levels to front, ordering by first appearance, reversing, and randomly shuffling), and tools for modifying factor levels (including collapsing rare levels into other, 'anonymising', and manually 'recoding').

https://github.com/tidyverse/forcats
https://forcats.tidyverse.org/
https://tidyverse.r-universe.dev/forcats
https://cran.r-project.org/package=forcats

```{r}
library(forcats)
```



### BeeBDC

BeeBDC: Occurrence Data Cleaning
Flags and checks occurrence data that are in Darwin Core format. The package includes generic functions and data as well as some that are specific to bees. This package is meant to build upon and be complimentary to other excellent occurrence cleaning packages, including 'bdc' and 'CoordinateCleaner'. This package uses datasets from several sources and particularly from the Discover Life Website, created by Ascher and Pickering (2020). For further information, please see the original publication and package website. Publication - Dorey et al. (2023) <doi:10.1101/2023.06.30.547152> and package website - Dorey et al. (2023) <https://github.com/jbdorey/BeeBDC>.

https://jbdorey.github.io/BeeBDC/


```{r}
library(BeeBDC)
```


### dbc

dbc: Dictionary-Based Cleaning

Tools for dictionary-based data cleaning.

https://epicentre-msf.r-universe.dev/dbc
https://github.com/epicentre-msf/dbc
NOT ON CRAN

```{r}
library(dbc)
```




### dataspice

dataspice: Create Lightweight Schema.org Descriptions of Data

The goal of 'dataspice' is to make it easier for researchers to create basic, lightweight, and concise metadata files for their datasets. These basic files can then be used to make useful information available during analysis, create a helpful dataset "README" webpage, and produce more complex metadata formats to aid dataset discovery. Metadata fields are based on the 'Schema.org' and 'Ecological Metadata Language' standards.

https://docs.ropensci.org/dataspice
https://ropensci.r-universe.dev/dataspice
https://github.com/ropensci/dataspice
https://cran.r-project.org/package=dataspice

```{r}
library(dataspice)
```




### labelmachine

labelmachine: Make Labeling of R Data Sets Easy

Assign meaningful labels to data frame columns. 'labelmachine' manages your label assignment rules in 'yaml' files and makes it easy to use the same labels in multiple projects.

https://a-maldet.r-universe.dev/labelmachine
https://a-maldet.github.io/labelmachine
https://cran.r-project.org/package=labelmachine
https://github.com/a-maldet/labelmachine


```{r}
library(labelmachine)
```


### libr

libr: Libraries, Data Dictionaries, and a Data Step for R
https://dbosak01.r-universe.dev/libr
Contains a set of functions to create data libraries, generate data dictionaries, and simulate a data step. The libname() function will load a directory of data into a library in one line of code. The dictionary() function will generate data dictionaries for individual data frames or an entire library. And the datestep() function will perform row-by-row data processing.




```{r}
library(libr)
```



### fctutils

fctutils: Advanced Factor Manipulation Utilities
https://guokai8.r-universe.dev/fctutils
Provides a collection of utility functions for manipulating and analyzing factor vectors in R. It offers tools for filtering, splitting, combining, and reordering factor levels based on various criteria. The package is designed to simplify common tasks in categorical data analysis, making it easier to work with factors in a flexible and efficient manner.




```{r}
library(fctutils)
```


### Amelia

Amelia: A Program for Missing Data

A tool that "multiply imputes" missing data in a single cross-section (such as a survey), from a time series (like variables collected for each year in a country), or from a time-series-cross-sectional data set (such as collected by years for each of several countries). Amelia II implements our bootstrapping-based algorithm that gives essentially the same answers as the standard IP or EMis approaches, is usually considerably faster than existing approaches and can handle many more variables. Unlike Amelia I and other statistically rigorous imputation software, it virtually never crashes (but please let us know if you find to the contrary!). The program also generalizes existing approaches by allowing for trends in time series across observations within a cross-sectional unit, as well as priors that allow experts to incorporate beliefs they have about the values of missing cells in their data. Amelia II also includes useful diagnostics of the fit of multiple imputation models. The program works from the R command line or via a graphical user interface that does not require users to know R.

https://matthewblackwell.r-universe.dev/Amelia


```{r}
library(Amelia)
```




### inspectdf

inspectdf: Inspection, Comparison and Visualisation of Data Frames

A collection of utilities for columnwise summary, comparison and visualisation of data frames. Functions report missingness, categorical levels, numeric distribution, correlation, column types and memory usage.

https://alastairrushworth.r-universe.dev/inspectdf

```{r}
library(inspectdf)
```




### esquisse
esquisse
https://dreamrs.github.io/esquisse/

```{r}
library(esquisse)
```



### surveyexplorer


surveyexplorer: Quickly Explore Complex Survey Data

Visualize and tabulate single-choice, multiple-choice, matrix-style questions from survey data. Includes ability to group cross-tabulations, frequency distributions, and plots by categorical variables and to integrate survey weights. Ideal for quickly uncovering descriptive patterns in survey data.

https://cran.r-project.org/web/packages/surveyexplorer/index.html
https://github.com/liamhaller/surveyexplorer
https://cran.r-project.org/web/packages/surveyexplorer/surveyexplorer.pdf
https://liamhaller.r-universe.dev/surveyexplorer

```{r}
library(surveyexplorer)
```

### gtsummary

gtsummary: Presentation-Ready Data Summary and Analytic Result Tables

Creates presentation-ready tables summarizing data sets, regression models, and more. The code to create the tables is concise and highly customizable. Data frames can be summarized with any function, e.g. mean(), median(), even user-written functions. Regression models are summarized and include the reference rows for categorical variables. Common regression models, such as logistic regression and Cox proportional hazards regression, are automatically identified and the tables are pre-filled with appropriate column headers.

https://cran.r-project.org/web/packages/gtsummary/index.html
https://www.danieldsjoberg.com/gtsummary/
https://github.com/ddsjoberg/gtsummary/
https://ddsjoberg.r-universe.dev/gtsummary

```{r}
library(gtsummary)
```


### descriptr

descriptr: Generate Descriptive Statistics

Generate descriptive statistics such as measures of location, dispersion, frequency tables, cross tables, group summaries and multiple one/two way tables.

https://rsquaredacademy.r-universe.dev/descriptr
https://descriptr.rsquaredacademy.com/
https://cran.r-project.org/package=descriptr
https://github.com/rsquaredacademy/descriptr/

ds_summary_stats from descriptr

```{r}
library(Desctable)
```


### Desctable

desctable: Produce Descriptive and Comparative Tables Easily

Easily create descriptive and comparative tables. It makes use and integrates directly with the tidyverse family of packages, and pipes. Tables are produced as (nested) dataframes for easy manipulation.

https://desctable.github.io/index.html
https://cran.r-project.org/package=desctable
https://github.com/desctable/desctable/
https://desctable.r-universe.dev/desctable

```{r}
install.packages("desctable")
library(Desctable)
```



### stargazer
https://cloud.r-project.org/web/packages/stargazer/index.html

```{r}
library(stargazer)
```


### psych
describe and describeBy, from the psych package

```{r}
library(psych)
```


### desctable

install.packages("desctable")

```{r}
library(desctable)
```


### pastecs

stat.desc, from the pastecs package

```{r}
library(pastecs)
```


### nomclust

nomclust: Hierarchical Cluster Analysis of Nominal Data
Similarity measures for hierarchical clustering of objects characterized by nominal (categorical) variables. Evaluation criteria for nominal data clustering.

https://zdeneksulc.r-universe.dev/nomclust
https://cran.r-project.org/package=nomclust

### ordinal

ordinal: Regression Models for Ordinal Data
Implementation of cumulative link (mixed) models also known as ordered regression models, proportional odds models, proportional hazards models for grouped survival times and ordered logit/probit/... models. Estimation is via maximum likelihood and mixed models are fitted with the Laplace approximation and adaptive Gauss-Hermite quadrature. Multiple random effect terms are allowed and they may be nested, crossed or partially nested/crossed. Restrictions of symmetry and equidistance can be imposed on the thresholds (cut-points/intercepts). Standard model methods are available (summary, anova, drop-methods, step, confint, predict etc.) in addition to profile methods and slice methods for visualizing the likelihood function and checking convergence.

https://runehaubo.r-universe.dev/ordinal
https://cran.r-project.org/package=ordinal
https://github.com/runehaubo/ordinal

```{r}

```



### classInt

classInt: Choose Univariate Class Intervals

Selected commonly used methods for choosing univariate class intervals for mapping or other graphics purposes.

https://r-spatial.r-universe.dev/classInt
https://r-spatial.github.io/classInt/
https://cran.r-project.org/package=classInt
https://github.com/r-spatial/classint

```{r}

```


#### Examples


##### Settlement Hierarchy Levels

Classifying the hierarchical levels of settlement systems from archaeological data can be notoriously arbitrary without more contextual information. The main issues are why one set of size breaks is better than another, how many levels to include, and the temptation to ideosyncratic/unsystematic/non-reproducible human choices that vary between periods/cases/regions.

##### Number of Classes for Each Period

To get around these issues, I employ a systematic and reproducible two-part methodology. First, the number of hierarchical levels is chosen as the rounded mean of several different established methods of computing the **number of classes** (as opposed to partitioning the class intervals themselves) for a univariate distribution such as settlement population sizes. The methods that performed the best on our right-skew city size distributions are:

  1. **Sturges' formula** where bin sizes are based on the range of the data [@Sturges1926], calculated using the "grDevices::nclass" function [@R2022]
  2. **Scott's method** based on the standard error of a Gaussian distribution fitted to the data [@Scott1992]. To make estimates of this method valid, Log Population was used instead of population. Scott's classes were calculated using the "grDevices::nclass" function [@R2022]
  3. **Jiang's Head/Tail method**, designed for heavy-tailed distributions, partitioning the distribution by the mean until the head-tail ratio is deskewed beyond a threshold level [@Jiang2013]. Jiang's method was calculated using the "classInt" R package [@Bivand2022].
  4. **Manually** via the traditional archaeological method of locating gaps in the distribution through close visual inspection of histograms and ECDFs (and a mind towards contextually-likely settlement hierarchy size classes). These are as follows:
  
```{r, 'Manual Method Settlement Hierarchy Classes', echo=FALSE,message=FALSE,warning=FALSE}
m = All_AggPoly@data %>% 
          group_by(PeriodNum) %>% 
          summarise(
              MinPop = min(Population.s2, na.rm=T),
              MaxPop = max(Population.s2, na.rm=T))

m$brks <- rev(c(", 250, 750, 2000, 4000, 8000, ",
            ", 200, 650, 2000, 10000, 20000, ",
            ", 150, 600, 1250, 3000, 8000, ",
            ", 100, 500, 2000, 7000, 12500, ",
            ", 200, 500, 1250, 2500, 7500, ",
            ", 100, 300, 600, 2000, ",
            ", 125, 350, 1250, 3000, 7500, ",
            ", 300, 2500, ",
            ", 100, 300, 600, 1500, ",
            ", 100, 200, 400, ",
            ", 250, 1000, 2000, 5000, ",
            ", 200, 1000, 2500, 6000, ",
            ", 100, 800, 3500, 7000, ",
            ", 150, 400, 800, ",
            ", 100, 300, 1000, 3000, ",
            ", 120, ",
            ", 200, "))

ManualHierClasses <- data.frame(
        PeriodNum = as.character(c(1:17)),
        Period = c("EF", "EF_MF", "MF", "MF_LF", "LF", "LF_TF", "TF", "TF_CL", "CL", "CL_ET", "ET", "ET_LTAzI", "LTAzI", "LTAzI_EA", "EA", "EA_LA", "LA"),
        Breaks = paste0(round(m$MinPop-1,0),m$brks,round(m$MaxPop+1,0)),
        NumClasses = c(2,2,5,4,5,5,5,4,5,3,6,5,6,6,6,6,6))

knitr::kable(ManualHierClasses, "pipe")

rm(m)
```
  
All four of these were performed on the population data of each period, and the row-wise mean and median (rounded to the nearest integer) were calculated.

```{r, 'Number of Settlement Hierarchy Classes', message=FALSE,warning=FALSE}
Manual <- ManualHierClasses$NumClasses
Sturges <- NA
Scott <- NA
HeadTails <- NA

for (i in 1:length(Poly_List)){
  Sturges[i] <- nclass.Sturges(Poly_List[[i]]$Population.s2)
  Scott[i] <- nclass.scott(Poly_List[[i]]$Log_Population.s2)
  HTbreaks <- classIntervals(Poly_List[[i]]$Population.s2, style = "headtails")
  HeadTails[i] <- length(HTbreaks$brks) - 1
}

levels <- data.frame(PeriodNum=ManualHierClasses$PeriodNum, 
                     Period=ManualHierClasses$Period, 
                     Sturges=Sturges, Scott=Scott, 
                     HeadTails=HeadTails, Manual=Manual)

levels <- levels %>% rowwise() %>% mutate(
  Median = round(median(c(Sturges,HeadTails,Manual)),0),
  Mean = round(mean(c(Sturges,HeadTails,Manual)),0),
) %>% ungroup()

knitr::kable(levels, "pipe" ,align="c")

rm(Manual, Sturges, Scott, HeadTails, HTbreaks)
```

Surprisingly, the _median_ of the four methods is **_very_** close to the "Manual" archaeological reasoning method undertaken by myself -- only differing by 1 for a single period (otherwise identical). The difference between the mean and the median is also quite small. More specifically, the mean is higher than the median by 1 for four periods (otherwise identical).

The Heads/Tails method has a very low sensitivity to such changes overall by design. Because Heads/Tails splits progressively down the distribution by cutting at the mean of the left tail, the increasing number of classes only reflects the magnitude of right-skew (completely insensitive to the number + variability of cases in the left tail). By contrast, Sturges' and Scott's methods are especially responsive to the growing ranges of the distributions, estimating systematically much higher numbers of classes for periods with a large number+range of settlements in a wide-ranging hierarchy (and vice versa). As such, they are both very sensitive to changes in the distribution as a whole. 

In this sense, the mean and median central tendencies nicely balance the extremes of the three systematic methods. The near identity of the mean and median, and their close similarity to the independently estimated "Manual archaeological breaks reasoning" method, both reccomend this middling set of metrics. I think that the mean performs the best because it rounds upwards for periods where there is a high right-skew. Whereas my manual classifications were cautious not to over-split the settlements in the fat right tail, the higher mean seems to indicate that my right tail bin sizes were too wide (i.e. they deserved to be subdivided in some cases, which is being picked up in the mean because of the suystematic shifts in the other metrics). I have therefore chosen to use the mean as the number of hierarchical levels to classify for each period.


##### Classification Method

The ["classInt" R Package](https://cran.r-project.org/web/packages/classInt) [@Bivand2022] provides a number of functions for classifying the univariate settlement size (population) hierarchy into the number of groups specified above (i.e. the integer-rounded mean of the four methods). Of these methods, some are not appropriate for a right-skew distribution, and others don't enable us to specify the number of classes. The best performing methods are kmeans, fisher and jenks. All others leave almost the entire fat left tail of the distribution unpartitioned. 

  * kmeans
  * fisher
  * jenks
When the settlement size distribution is not heavily right-skew, all three methods perform about the same. However, with greater degrees of right-skew, Fisher's and Jenks' methods leave most of the left-tail unpartitioned. As seen below, only kmeans provides a partitioning method robust to higher levels of right-skew, creating size classes quite similar to the Manual Archy method. (Note that the intervals are calculated on unlogged data, while the x-axes are logged to make the groups clearer.) As such, Kmeans seems to provide a systematic means of partitioning the settlement distributions.

```{r, 'Hierarchy Classes Example 1', echo=FALSE, message=FALSE,warning=FALSE}
p=15
x=Poly_List[[p]]$Population.s2
nc=levels$Mean[p]
pal1=brewer.pal(nc,"Set2")
fb=as.numeric(unlist(strsplit(ManualHierClasses$Breaks[p], split=", ")))

kmbrks <- classIntervals(x, style = "kmeans", n=nc)
hcbrks <- classIntervals(x, style = "hclust", n=nc)
bcbrks <- classIntervals(x, style = "bclust", n=nc)
fbrks <- classIntervals(x, style = "fisher", n=nc)
jbrks <- classIntervals(x, style = "jenks", n=nc)
maxbrks <- classIntervals(x, style = "maximum", n=nc)
fixbrks <- classIntervals(x, style = "fixed", fixedBreaks=fb)

par(mfrow = c(2, 2))
plot(fbrks, pal = pal1, main = paste0("Fisher, ",levels$Period[p]," ",levels$PeriodNum[p]),log = 'x',xlim=c(min(x)-1, max(x)+5), xlab="Log Population", ylab="ECDF(Population)")
plot(jbrks, pal = pal1, main = paste0("Jenks, ",levels$Period[p]," ",levels$PeriodNum[p]),log = 'x',xlim=c(min(x)-1, max(x)+5), xlab="Log Population", ylab="ECDF(Population)")
plot(kmbrks, pal = pal1, main = paste0("Kmeans, ",levels$Period[p]," ",levels$PeriodNum[p]),log = 'x',xlim=c(min(x)-1, max(x)+5), xlab="Log Population", ylab="ECDF(Population)")
plot(fixbrks, pal = pal1, main = paste0("Manual, ",levels$Period[p]," ",levels$PeriodNum[p]),log = 'x',xlim=c(min(x)-1, max(x)+5), xlab="Log Population", ylab="ECDF(Population)")
par(mfrow = c(1, 1))

```

```{r, 'Hierarchy Classes Example 2', echo=FALSE, message=FALSE,warning=FALSE}
p=5
x=Poly_List[[p]]$Population.s2
nc=levels$Mean[p]
pal1=brewer.pal(nc,"Set2")
fb=as.numeric(unlist(strsplit(ManualHierClasses$Breaks[p], split=", ")))

kmbrks <- classIntervals(x, style = "kmeans", n=nc)
hcbrks <- classIntervals(x, style = "hclust", n=nc)
bcbrks <- classIntervals(x, style = "bclust", n=nc)
fbrks <- classIntervals(x, style = "fisher", n=nc)
jbrks <- classIntervals(x, style = "jenks", n=nc)
maxbrks <- classIntervals(x, style = "maximum", n=nc)
fixbrks <- classIntervals(x, style = "fixed", fixedBreaks=fb)

par(mfrow = c(2, 2))
plot(fbrks, pal = pal1, main = paste0("Fisher, ",levels$Period[p]," ",levels$PeriodNum[p]),log = 'x',xlim=c(min(x)-1, max(x)+5), xlab="Log Population", ylab="ECDF(Population)")
plot(jbrks, pal = pal1, main = paste0("Jenks, ",levels$Period[p]," ",levels$PeriodNum[p]),log = 'x',xlim=c(min(x)-1, max(x)+5), xlab="Log Population", ylab="ECDF(Population)")
plot(kmbrks, pal = pal1, main = paste0("Kmeans, ",levels$Period[p]," ",levels$PeriodNum[p]),log = 'x',xlim=c(min(x)-1, max(x)+5), xlab="Log Population", ylab="ECDF(Population)")
plot(fixbrks, pal = pal1, main = paste0("Manual, ",levels$Period[p]," ",levels$PeriodNum[p]),log = 'x',xlim=c(min(x)-1, max(x)+5), xlab="Log Population", ylab="ECDF(Population)")
par(mfrow = c(1, 1))

```

These two periods are also representative in terms of their mixed geographical distribution of size classes across the region, further cross-validating the Kmeans settlement hierarchy class intervals. 

##### Classifying All Periods

Now we can classify all periods using this method

```{r, 'Calculate Hierarchy Classes for all Periods', message=FALSE,warning=FALSE}

for (i in 1:length(Poly_List)){
  
  nc=levels$Mean[i]
  
  kmbrks <- classIntervals(Poly_List[[i]]$Population.s2, style = "kmeans", n=nc)
  
  Poly_List[[i]]$SetHierLevel <-  cut(Poly_List[[i]]$Population.s2, 
                                      kmbrks$brks, labels = FALSE, 
                                      include.lowest = TRUE)
}

```




## Data cleaning for statistical purpose: Software for cleaning data

Mark van der Loo
Edwin de Jonge

https://data-cleaning.r-universe.dev/packages

### validate

validate: Data Validation Infrastructure
Declare data validation rules and data quality indicators; confront data with them and analyze or visualize the results. The package supports rules that are per-field, in-record, cross-record or cross-dataset. Rules can be automatically analyzed for rule type and connectivity. Supports checks implied by an SDMX DSD file as well. See also Van der Loo and De Jonge (2018) <doi:10.1002/9781118897126>, Chapter 6 and the JSS paper (2021) <doi:10.18637/jss.v097.i10>.

https://data-cleaning.r-universe.dev/validate
https://cran.r-project.org/package=validate
https://github.com/data-cleaning/validate




### validatetools


### validatesuggest


### deductive


### deducorrect


### lintools


### errorlocate


### dcmodify


### editrules

https://data-cleaning.r-universe.dev/editrules
https://data-cleaning.r-universe.dev/dcmodify
https://data-cleaning.r-universe.dev/errorlocate
https://data-cleaning.r-universe.dev/lintools
https://data-cleaning.r-universe.dev/validatetools
https://data-cleaning.r-universe.dev/validatesuggest
https://data-cleaning.r-universe.dev/deductive
https://data-cleaning.r-universe.dev/deducorrect
https://markvanderloo.r-universe.dev/rspa

Mark van der Loo
Edwin de Jonge

https://data-cleaning.r-universe.dev/packages
