---
title: "Wide Format Conversion of TMP Databases"
author: "Rudolf Cesaretti"
date: "2025-02-27"
output: html_document
---

```{r}
rm(list = ls())

```


# Load Packages

```{r}
# Load the required package
library(tidyverse)
library(readxl)
library(stringi)

```


# Set Working Directory

```{r}
DIR = "C:/Users/TJ McMote/ASU Dropbox/Rudolf Cesaretti/GitHubRepos/TMP-database/"
```


# DF8 Wide 

Merge DF8 Tables for Raw (original numeric) Values Dataset


### Step 1: Import all excel files from the specified directory into a list of dataframes

```{r}

file_list <- list.files(
  path = paste0(DIR, "data/raw/DF8"),
  pattern = "\\.xlsx$", full.names = TRUE
)
file_names <- tools::file_path_sans_ext(basename(file_list))
data_list <- map(file_list, read_excel)
names(data_list) <- file_names

```

### Step 2: Rearrange the list so the last dataframe is first, then join all

```{r}

# Rearrange the list so the last dataframe is first, then join all
DF8_wide <- data_list %>%
  {c(tail(., 1), head(., -1))} %>%  # Reorder the list
  reduce(full_join, by = "SSN") %>% # Join all dataframes by SSN
  rename(`SITE_` = `Site.`,
         `UNIT_` = `Unit`,
         `DRAINS_ARCH` = `DRAINS.x`,
         `DRAINS_MODERN` = `DRAINS.y`) # Rename columns

```


### Step 3: Check for any discrepancies between the SITE and UNIT columns

```{r}

checksum <- DF8_wide %>% 
  dplyr::select(SSN, SITE, SITE_, UNIT, UNIT_) %>% 
  mutate(check_SITE = ifelse(SITE == SITE_, 1, 0),
         check_UNIT = ifelse(UNIT == UNIT_, 1, 0)) %>% 
  summarize(checksum_SITE = sum(check_SITE),
            checksum_UNIT = sum(check_UNIT))

if (checksum$checksum_SITE == nrow(DF8_wide) & checksum$checksum_UNIT == nrow(DF8_wide)) {
  print("CHECKS PASSED: SITE + SITE_ and UNIT + UNIT_ columns are identical across all rows")
} else {
  print("WARNING: SITE + SITE_ and UNIT + UNIT_ columns are NOT identical across all rows")
}

DF8_wide <- DF8_wide %>% 
    dplyr::select(-SITE, -UNIT_) %>% 
    rename(SITE = SITE_)
```

### Step 4: Misc Changes

```{r}
DF8_wide <- DF8_wide %>% 
  mutate(NORTHING = NORTHING*10,
         EASTING = EASTING*10,
         SITE = str_replace_all(SITE, " ", ""))
```


### Step 5: Save the DF8_wide dataframe to a CSV file

```{r}
write_csv(DF8_wide, paste0(DIR, "data/raw/DF8_wide.csv"))
```

### Clean Up

```{r}
rm(checksum, file_list, file_names, data_list)
```


# DF9 Wide

### Step 1: 

```{r}

```

### Step 2: 

```{r}

```

### Step 3: 

```{r}

```

### Step 4: 

```{r}

```

### Step 5: 

```{r}

```

### Step 6: 

```{r}

```

### Step 7: 

```{r}

```

### Step 8: 

```{r}

```

### Step 9: 

```{r}

```

### Step 10: 

```{r}

```

### Step 11: 

```{r}

```

### Step 12: 

```{r}

```


### Clean Up

```{r}
rm(file_list, file_names, data_list)
```




# DF10 Wide



### Step 1: Import all CSV files from the specified directory

```{r}

file_list <- list.files(
  path = paste0(DIR, "data/raw/DF10"),
  pattern = "\\.csv$", full.names = TRUE
)
file_names <- tools::file_path_sans_ext(basename(file_list))
data_list <- map(file_list, ~ read_csv(.x, locale = locale(encoding = "Latin1")))
names(data_list) <- file_names
list2env(data_list, envir = .GlobalEnv)

df_list <- list_modify(df_list, Data_fieldWorkers = NULL, Data_labAnalysts = NULL)

```





### Step 2: Format totalsTable

```{r}

totals <- totalsTable %>%
  mutate(
    NewCount = ifelse(Where == "M", -9999, Where),
    NewCount = ifelse(is.na(NewCount), Count, NewCount),
    NewCount = as.numeric(NewCount)
  ) %>% # Replace 0 with NA where "Where" == "M"
  select(-Where, -Count) %>% # Remove "Where" column
  pivot_wider( # Pivot wider by ID
    id_cols = SSN,
    names_from = Variable,
    values_from = NewCount,
    values_fill = 0
  ) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>%
  arrange(SSN) %>%
  mutate(across(everything(), ~ na_if(., -9999)))

```






### Step 3: Format interpTable

```{r}

interps <- interpTable %>%
  mutate(
    NewCode = ifelse(Where == "M", -1, Where), # Replace 0 with -1 where "Where" == "M"
    NewCode = ifelse(is.na(NewCode), Code, NewCode),
    NewCode = as.numeric(NewCode)
  ) %>%
  left_join(interpCodes, by = join_by("NewCode" == "Code")) %>% # Join with interpCodes
  select(-Code, -Where) %>% # Remove unneeded columnws
  pivot_wider(id_cols = SSN, names_from = Variable, values_from = Description) %>% # Pivot wider by ID
  mutate(SSN = as.numeric(SSN)) %>%
  arrange(SSN)

```





### Step 4: Format codeTable

```{r}

codes <- codeTable %>%
  mutate(
    NewCode = ifelse(Where == "M", -1, Where), # Replace 0 with -1 where "Where" == "M"
    NewCode = ifelse(is.na(NewCode), Code, NewCode),
    NewCode = as.numeric(NewCode)
  ) %>%
  select(-Code, -Where) %>% # Remove unneeded columnws
  left_join(codeCodes, by = join_by("NewCode" == "Code")) %>% # Join with interpCodes
  pivot_wider(
    id_cols = SSN, names_from = Variable,
    values_from = Description, values_fill = "absent"
  ) %>% # Pivot wider by SSN
  mutate(SSN = as.numeric(SSN)) %>%
  arrange(SSN)

```





### Step 5: Format artifactTable

```{r}

artifacts <- artifactTable %>%
  mutate(
    NewCount = ifelse(Where == "M", -9999, Where), # Replace 0 with NA where "Where" == "M"
    NewCount = ifelse(is.na(NewCount), Count, NewCount),
    NewCount = as.numeric(NewCount)
  ) %>%
  select(-Where, -Count) %>%
  left_join(artifactCodes, by = c("ArtCode1" = "Code")) %>%
  rename(ArtCode4 = Description) %>%
  left_join(artifactCodes, by = c("ArtCode2" = "Code")) %>%
  rename(ArtCode5 = Description) %>%
  left_join(artifactCodes, by = c("ArtCode3" = "Code")) %>%
  rename(ArtCode6 = Description) %>%
  mutate(
    ArtCode6 = ifelse(ArtCode2 == 12 & ArtCode3 == 46, "figAzte", ArtCode6),
    ArtCode7 = paste(ArtCode4, ArtCode5, ArtCode6, sep = "_")#################################
  ) %>%
  select(SSN, ArtCode6, NewCount) %>%##################################
  pivot_wider(names_from = ArtCode6, values_from = NewCount, values_fill = 0) %>% # Pivot wider by SSN
  arrange(SSN) %>%
  mutate(across(everything(), ~ na_if(., -9999)))


```


### Step 6: Join formatted tables to provTable

```{r}

DF10_wide <- provTable %>%
  left_join(interps, by = "SSN") %>%
  left_join(codes, by = "SSN") %>%
  left_join(artifacts, by = "SSN") %>%
  left_join(totals, by = "SSN")

```



### Step 7 : Misc Changes

```{r}

DF10_wide <- DF10_wide %>%
  #rename(ceramicAnalyst1 = workLab1,
  #       ceramicAnalyst2 = workLab2,
  #       ceramicAnalyst3 = workLab3,
  #       fieldWorker1 = workFie1,
  #       fieldWorker2 = workFie2,
  #       fieldWorker3 = workFie3,
  #       fieldWorker4 = workFie4,
  #       fieldWorker5 = workFie5,
  #       Ceramic_cerOther_totCuan = Ceramic_cerOther_Cuan,
  #       Ceramic_cerOther_totMaza = Ceramic_cerOther_Maza,
  #       Ceramic_cerOther_totAzte = Ceramic_cerOther_Azte,
  #       Ceramic_cerOther_totPreCl = Ceramic_cerOther_PreCl,
  #       Ceramic_cerOther_totTezo = Ceramic_cerOther_Tezo,
  #       Ceramic_cerOther_totCoyo = Ceramic_cerOther_Coyotlatelco) %>% 
  mutate(Northing = Northing*10,
         Easting = Easting*10,
         Site = str_replace_all(Site, " ", ""))

```


### Step 8: Export the wide-format table as "DF10_wide.csv"

```{r}

write.csv(DF10_wide, paste0(DIR, "data/raw/DF10_wide.csv"), fileEncoding = "Latin1")

write.csv(archToSSN, paste0(DIR, "data/raw/DF10_archToSSN.csv"), fileEncoding = "Latin1")

```



### Clean Up

```{r}
rm(file_list, file_names, data_list)
```





Make sure DF8 has NA values where appropriate

Add character string variables to DF8
--use DF9 descriptions for DF8 



Variable Modifications
--Field workers to boolean
--Lab analysts to boolean




Standardize Variable Names

variable names concordance
upload table with var name concordance and use this as the means for equivalence
Add a new column to this table with the new standardized name
then use this table to rename the columns in the wide format tables

Reorder Columns based on new finalized variable ordering 



Standardized variable factor levels ------ did anything change??




